% !TEX program = pdflatex
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{3} % number sections, subsections, subsubsections
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% \IfFileExists{microtype.sty}{% use microtype if available
%   \usepackage[]{microtype}
%   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
% }{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\let\oldhyperref\hyperref
\renewcommand{\hyperref}[2][]{\oldhyperref[#1]{\underline{#2}}}

\author{}
\date{}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\textbf{Going from 1+1=2 to Quantum Mechanics}}
\fancyhead[R]{Chaidhat Chaimongkol}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\providecommand{\ket}[1]{\left\lvert #1 \right\rangle}
\providecommand{\bra}[1]{\left\langle #1 \right\rvert}
\providecommand{\braket}[1]{\left\langle #1 \right\rangle}
\providecommand{\Rarr}{\Rightarrow}
\providecommand{\Larr}{\Leftarrow}
\providecommand{\lrarr}{\Leftrightarrow}
\providecommand{\rarr}{\rightarrow}
\providecommand{\larr}{\leftarrow}
\providecommand{\lrar}{\leftrightarrow}
\providecommand{\Z}{\mathbb{Z}}
\providecommand{\R}{\mathbb{R}}
\providecommand{\N}{\mathbb{N}}
\providecommand{\C}{\mathbb{C}}
\providecommand{\Q}{\mathbb{Q}}
\providecommand{\infin}{\infty}

\title{Going from 1+1=2 to Quantum Mechanics}
\author{Chaidhat Chaimongkol}
\date{\today}

\begin{document}
\maketitle


Hi, my name is Chai and I'm taking an undergraduate quantum mechanics course at UCLA.
\bigskip

It's \underline{\href{https://www.reddit.com/r/ucla/comments/1lhbq9z/ece_128_hardest_class_at_ucla/}{known as one of the hardest classes at UCLA}.}
\bigskip

This text is written so that, at the start, you only need to know how to do basic middle school arithmetic.
\bigskip

In the end, you'll know how to do quantum mechanics problems.
\bigskip

This text is loosely sorted by complexity. Everything is derived from concepts introduced in previous sections. As we introduce mathematical concepts, we will use them to better understand quantum mechanics.
\bigskip

This is based on the Electrical and Computer Engineering 128 course
taught at the University of California, Los Angeles by Dr.~Leo Zhou.

License: CC0 No rights reserved



\section*{None of this is written by
	AI}\label{none-of-this-is-written-by-ai}

None of this is written by AI --- I am writing this by hand because I
find explaining concepts makes me understand them better too.

\section*{Everything must be defined
	above}\label{everything-must-be-defined-above}

Any mathematical term MUST be derived from first principles (or like
elementary school math) in a section before it.

Example

\hyperref[invariant-subspaces]{invariant
	subspace} →
\hyperref[subspaces]{subspace}
→
\hyperref[subsets]{subset}
→
\hyperref[set-theory]{set
	theory}


\section*{chaimongkol@ucla.edu}

Please let me know at
\href{mailto:chaimongkol@ucla.edu}{\nolinkurl{chaimongkol@ucla.edu}} on
how I'm doing or that you've even read this message. If I see support,
it makes me want to continue this series! The series is currently
incomplete.

\newpage
\makeatletter
\renewcommand{\l@subsection}{\@dottedtocline{2}{1.5em}{2.8em}}
\makeatother
\tableofcontents
\newpage

\section{Let}\label{let}

Ok when I say ``let,'' i'm just defining something to be another thing.

Example:

\begin{gather*}
	\text{Let blorple be blue}
\end{gather*}

Now if i say ``the sky is blorple'' then i actually mean that it is
blue.


\section{Variable, Addition}\label{variables-addition}

Let a variable be just a symbol that represents a literal thing.

I can define what number it represents by saying ``it's equal to this''

Example:

Let

\begin{gather*}
	x=1
\end{gather*}

Now if I add something to that variable, I'm going to write a double
arrow ``\(\Rarr\)'' on the left of it to show that ``Therefore this is
true''

Example:

\begin{gather*}
	x=1\\
	\Rarr x+1=1+1\\
	\Rarr x+1=2
\end{gather*}



\(\Rarr x+5=6\), as \(x=1\), we'd just put 1 in place of the \(x\)


\section{Matrix}\label{matrices}

Let a ``matrix'' be a set of numbers ordered in a neat grid. Now we
can manipulate a lot of numbers simultaneously. Interestingly enough,
they are the solution to a lot of real world including machine learning
and quantum mechanics.

Let's make a new
\hyperref[variables-addition]{variable}
called \(A\).

Let \(A\) be a matrix of size \(m\times n\) where \(m\) is the height of
the matrix (how many rows it has) and \(n\) is the width (how many
columns it has).

Example:

\begin{gather*}
	A=\begin{pmatrix}
		1 & 2 & 5 \\
		3 & 4 & 9
	\end{pmatrix}
\end{gather*}



See how \(A\) is also a variable? In this case, \(m=2, n=3\) as there
are 2 rows and 3 columns.

We use matrices to represent multiple numbers because they have an
equivalent in the real world. In this case, quantum states.

\section{Ket (State)}\label{ket-states}

Let \(\ket{x}\) (ket x) be a \(N \times 1\) \hyperref[matrices]{matrix} where \(N\) is any
positive integer (number with no decimals and is greater than 0).

Just like how \(x\) was a \hyperref[variable]{variable} used to represent ONE number,
\(\ket{x}\) can be used to represent any \(N\times 1\) matrix.

Example:

\begin{gather*}
	\ket{x}=\begin{pmatrix}
		1 \\
		2 \\
		7
	\end{pmatrix}\quad \ket{\psi}=\begin{pmatrix}
		9 \\
		9 \\
		0 \\
		0 \\
		0
	\end{pmatrix}\quad\ket{\phi}=\begin{pmatrix}
		0 \\
		1
	\end{pmatrix}
\end{gather*}



Note that \(\psi\) (psi) is pronounced sye and \(\phi\) (phi) is
pronounced fye. They are just like \(x\) but idk why quantum physicists
love greek characters.

\section{Photon, Wave-particle
  Duality}\label{photons-wave-particle-duality}

Let this be a photon. It's like a particle of light.


People, like Newton, favored the idea that light was a particle. This
was until Young, a person, performed an experiment called the double
slit experiment which showed light interacting like a wave. Maxwell,
another person, showed that light is an electromagnetic wave --- an
oscillation of the electric and magnetic fields themselves. Einstein,
another person, then explained the photoelectric effect, showing light
also behaves as a particle, so light has both wave AND particle
character.

Present day, the majority of people concur that light is a wave and
particle simultaneously.

This means that photons can have a ``state.'' These can be represented
as \hyperref[ket-states]{kets}!

Sometimes we observe photons having a ``spin'' --- it's not really
spinning, but physicists use the word for an intrinsic angular momentum
that particles carry. Photons have spin 1; for photons this shows up
as polarization (e.g. circular polarization corresponds to the two
helicity states).

\section{Superposition, Matrix
  Addition}\label{superpositions-matrix-addition}

A photon can have a probability of \(x\%\) of spinning up (+). This
means it would have \((100-x)\%\) of spinning down (-), since the
probabilities must add up. The photon actually is both spinning up and
spinning down until we observe it, where it collapses to either spinning
up or spinning down. Schröndinger's cat.


We call this position of spinning both up and down simultaneously a
``superposition.''

Let's just say the \hyperref[ket-states]{kets} \(\ket{+}\) is the state of spinning up and \(\ket{-}\)
is the state of spinning down.

Let a ``state'' just be a \(N\times 1\) matrix like what we just said
above. So \(\ket{x},\ket{\psi},\ket{\phi}\) are all states.

Some states can be made of other states.

Example:

Let

\begin{gather*}
	\ket{x}=\begin{pmatrix}
		1 \\
		0
	\end{pmatrix}\quad
	\ket{y}=\begin{pmatrix}
		0 \\
		1
	\end{pmatrix}
\end{gather*}


If i multiply a matrix by a number (let this be called a ``scalar''), we
apply the multiplication across all values.

\begin{gather*}
	\Rarr 3\ket{x}=3\begin{pmatrix}1\\0\end{pmatrix}=\begin{pmatrix}3\times1\\3\times0\end{pmatrix}=\begin{pmatrix}3\\ 0\end{pmatrix}
\end{gather*}


If we add \(\ket{x}+3\ket{y}\) we get

\begin{gather*}
	\Rarr\ket{x}+3\ket{y}=\begin{pmatrix}
		1 \\
		0
	\end{pmatrix}+3
	\begin{pmatrix}
		0 \\
		1
	\end{pmatrix}
	=\begin{pmatrix}
		1 \\
		3
	\end{pmatrix}
\end{gather*}


Notice how we're adding the \(n\)-th elements of the states together for
each row?

We can represent these two states coexisting (i.e., a superposition)
using math now!

\begin{gather*}
	\ket{\psi}=\alpha\ket{+}+\beta\ket{-}
\end{gather*}

where \(\alpha,\beta\) are just numbers. See how they are a combination
of up and down states?

There is \(|\alpha|^2\) probabilty that the photon, once observed, spins
upwards and \(|\beta|^2\) probability it spins downwards once observed.



\(|a|\) is just a way of saying make \(a\) positive if it is negative.

Example:

\begin{gather*}
	|(-3)|=3\\
	|5|=5
\end{gather*}

Note that

\begin{gather*}
	|\alpha|^2+|\beta|^2=1
\end{gather*}

as you can't have more than 100\% or less than 0\% probability of
something existing.

I'm going to write a ton of mathematical definitions which will build up
how to do the physics in the next section!


\section{Set Theory}\label{set-theory}

Ok this is just for notation sake but i want to make my life a bit
easier.

\subsection{Element}\label{element}

an element is either a variable or a number or a literal in our
definition

\begin{gather*}
	1,5,a,\text{cat},...
\end{gather*}

\subsection{Set}\label{set}

Let a set be an unordered collection of
\hyperref[element]{elements}

Example:

This is a set \(S\)

\begin{gather*}
	S=\{3,5,1982549\}\\
	S_2=\{2n\quad|\quad n=1,2,5\}=\{2,4,10\}
\end{gather*}

The \textbar{} means ``such that''

\subsection{Subset}\label{subsets}

\hyperref[set]{sets}
can be subsets (or equal) of another set if all \hyperref[element]{elements} of a
\hyperref[set]{set}
is in another
\hyperref[set]{set}.

Example:

\begin{gather*}
	S\subset\{3,5,1982549,67\}
\end{gather*}

We have a line underneath to show it could be a subset OR equal

\begin{gather*}
	S\subseteq\{3,5,1982549\}
\end{gather*}

\subsection{In}\label{in}

We can say an
\hyperref[element]{element}
is ``in'' a
\hyperref[set]{set}
if it's in a set (duh)

Example:

\begin{gather*}
	3\in S\\
	4\notin S
\end{gather*}
\subsection{Union}\label{union}

The union of two sets \(A\) and \(B\) is the set of all elements that are in \(A\) or \(B\) or both.

\[
	A\cup B=\{x\in A \text{ or } x\in B\}
\]

\subsection{Intersection}\label{intersection}

The intersection of two sets \(A\) and \(B\) is the set of all elements that are in \(A\) and \(B\).

\[
	A\cap B=\{x\in A \text{ and } x\in B\}
\]

\subsection{Standard Number System}\label{standard-number-systems}

There are some standard number systems we gotta keep in mind

\begin{gather*}
	\{1,2,3,...\}\subseteq\mathbb{N}\\
	\{...,-3,-2,-1,0,1,2,3,...\}\subseteq\mathbb{Z}\\
	\{\frac{p}{q}\quad\forall p\in \mathbb{Z},q\neq 0 \}\subseteq\mathbb{Q}\\
	\{\sqrt{2},\pi,...\}\in \mathbb{R}\\
	\{3+1i, 5+5i, 2+0i\}\in\mathbb{C}=\mathbb{F}
\end{gather*}

Where

\begin{itemize}
	\tightlist
	\item
	      \(\mathbb{N}\) are natural numbers
	\item
	      \(\mathbb{Z}\) are integers
	\item
	      \(\mathbb{R}\) are real numbers
	\item
	      \(\mathbb{C}\) are \hyperref[complex-numbers]{complex numbers}, \(\mathbb{F}\) are fields
\end{itemize}

where \(\forall p\in \mathbb{Z},q\neq 0\) means for all integers \(p\)
AND \(q\neq 0\)

\subsection{Partitions}\label{partitions}
A partition of a set \(S=\{C_1,C_2,…,C_N\}\) is a way of splitting \(S\) into
\begin{enumerate}
	\item
	      disjoint subsets
	      \[
		      C_\alpha  \cap C_\beta = \emptyset \quad \forall \quad \alpha \neq \beta
	      \]
	\item
	      that cover \(S\)
	      \[
		      \bigcup_{\alpha=1}^N C_\alpha = S
	      \]
\end{enumerate}
\subsection{Strings}\label{strings}
A string is an ordered sequence of symbols from a finite alphabet. The alphabet is denoted with the set notation (a set of possible symbols to use as an alphabet). Length is specified as the superscript, but if it is unlimited, then it is denoted with an asterisk $*$.

Example
\[
	01101011 \in \{0,1\}^*
\]
\[
	cba \in \{a,b,c\}^3
\]
\[
	++++-- \notin \{+,-\}^2
\]


\section{Function}\label{functions}

Also let us notate a state as a function of time. What's a function?

Example:

if i let \(f\) be a function that accepts \(x\) as an argument, like
this

\begin{gather*}
	f(x)=4x
\end{gather*}

then

\begin{gather*}
	f(5)=4(5)=20
\end{gather*}

A function \(f\), can also be thought of mapping numbers from a set \(A\) to set \(B\). It's
written like this

\begin{gather*}
	f:A\rarr B
\end{gather*}

A map just links a thing to another thing.

Example

Let \(q\) be a function

\begin{gather*}
	q:\Z\rarr\Z\\
	q(x)=3x
\end{gather*}

q essentially maps all integers to another integer, which is three times
itself.

\begin{gather*}
	q(67)=201
\end{gather*}

\section{Summation}\label{summation}
This is a function that adds up a series of numbers.
\[
	\sum_{i=1}^n a_i \triangleq a_1 + a_2 + \cdots + a_n
\]

\section{Trigonometry}\label{trigonometry}


There's these \hyperref[functions]{functions} that are useful to know about.
\[
	\sin(\theta), \cos(\theta), \tan(\theta)
\]
Where \(\theta\) is the angle in radians.

1 radians is \(\frac{180}{\pi}\) degrees. The radian to degree conversion is linear.

For the purpose of this course, we only need to think about trigonometry as functions. No need to mess around with triangles, but it sure will be helpful to know more about them.

Also know these fun facts

\[
	\sin(\alpha\pm \beta) = \sin\alpha\cos\beta \pm \cos\alpha\sin\beta
\]
\[
	\cos(\alpha\pm \beta) = \cos\alpha\cos\beta \mp \sin\alpha\sin\beta
\]
\[
	\tan(\alpha\pm \beta) = \frac{\tan\alpha \pm \tan\beta}{1 \mp \tan\alpha\tan\beta}
\]

This is called the Pythagorean identity
\[
	\sin^2(\theta) + \cos^2(\theta) = 1
\]

\section{Calculus}\label{calculus}
\subsection{Limit}\label{limits}
Let a "limit" be a function that slowly moves something towards a value. For example, if \(h\) approaches \(\infin\), the equation below evaluates to a value. It's helpful when if we just simply set \(h=0\) or \(h=\infin\), then we divide by zero or for some reason can't compute \(h\).

\[
	\lim_{h\to \infin} 3+\frac{5}{h}=3
\]

\subsection{Differentiation}\label{differentiation}

To differentiate something means to see how fast it changes with respect
to (w.r.t) something else changing, for example, time \(t\). Imagine you step forward in time by a very small step \(h\).
As \(h\rarr 0\), the change in \(x\) becomes the derivative of \(x\) with respect to time.

\[
	\frac{d}{dt}f(t)=\lim_{h\to 0} \frac{f(t+h)-f(t)}{h}
\]

To do this, us humans developed some nifty tricks. Here's all of them.

\begin{align*}
	\frac{d}{dx}[f(x) \pm g(x)]                & = f'(x) \pm g'(x)                                                    \\
	\frac{d}{dx}[x^n]                          & = nx^{n-1}                                                           \\
	\frac{d}{dx}[f(x)g(x)]                     & = f'(x)g(x) + f(x)g'(x)                                              \\
	\frac{d}{dx}[f(g(x))]                      & = f'(g(x))\,g'(x)                                                    \\
	\frac{d}{dx}[e^x]                          & = e^x                                                                \\
	\frac{d}{dx}[a^x]                          & = a^x \ln(a)                                                         \\
	\frac{d}{dx}[\ln(x)]                       & = \frac{1}{x}                                                        \\
	\frac{d}{dx}[\log_a(x)]                    & = \frac{1}{x \ln(a)}                                                 \\
	\frac{d}{dx}[\sin(x)]                      & = \cos(x)                                                            \\
	\frac{d}{dx}[\cos(x)]                      & = -\sin(x)                                                           \\
	\frac{d}{dx}[\tan(x)]                      & = \left(\frac{1}{\cos(x)}\right)^2                                   \\
	\frac{d}{dx}\left(\frac{1}{\tan(x)}\right) & = \frac{1}{\sin^2 (x)}                                               \\
	\frac{d}{dx}\left(\frac{1}{\cos(x)}\right) & = \frac{\sin x}{\cos^2 x}                                            \\
	\frac{d}{dx}\left(\frac{1}{\sin(x)}\right) & = -\frac{\cos x}{\sin^2 x}                                           \\
	\frac{d}{dx}[\arcsin(x)]                   & = \frac{1}{\sqrt{1 - x^2}}                                           \\
	\frac{d}{dx}[\arccos(x)]                   & = -\frac{1}{\sqrt{1 - x^2}}                                          \\
	\frac{d}{dx}[\arctan(x)]                   & = \frac{1}{1 + x^2}                                                  \\
	\frac{d}{dx}[\operatorname{arccot}(x)]     & = -\frac{1}{1 + x^2}                                                 \\
	\frac{d}{dx}[\operatorname{arcsec}(x)]     & = \frac{1}{|x|\sqrt{x^2 - 1}}                                        \\
	\frac{d}{dx}[\operatorname{arccsc}(x)]     & = -\frac{1}{|x|\sqrt{x^2 - 1}}                                       \\
	\frac{d}{dx}[f(x)^{g(x)}]                  & = f(x)^{g(x)} \left[g'(x)\ln(f(x)) + \frac{g(x)\,f'(x)}{f(x)}\right] \\
	\frac{d}{dx}[f^{-1}(x)]                    & = \frac{1}{f'(f^{-1}(x))}                                            \\
	\frac{d^n}{dx^n}[f(x)g(x)]                 & = \sum_{k=0}^{n} \binom{n}{k} f^{(k)}(x)\,g^{(n-k)}(x)
\end{align*}

\subsection{Integration}\label{integration}
The opposite of \hyperref[differentiation]{differentiation} is integration. This is the Riemann sum.

\begin{gather*}
	\int_\alpha^\beta f(t)\cdot dt=\lim_{n\to\infin}\sum_{i=1}^n f(t_i)\cdot (t_i-t_{i-1})
\end{gather*}

Integration is much more hellish. You don't need to know integration for
this.

\subsection{Taylor Series}\label{taylor-series}
Taylor, a person, once said that

\[
	f(x) = \sum_{n=0}^{\infin}{\left(\frac{1}{n!}\right)\left(\frac{d^n f}{dx^n}\right)}|_a · (x - a)^n
\]

So...

\begin{gather*}
	e^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}+\frac{x^4}{4!}+...\\
	\cos x=1-\frac{x^2}{2!}+\frac{x^4}{4!}-\frac{x^6}{6!}+\cdots\\
	\sin x=x-\frac{x^3}{3!}+\frac{x^5}{5!}-\frac{x^7}{7!}+\cdots
\end{gather*}
\section{Complex Number}\label{complex-numbers}

Math is complicated. Some can say it's complex.

Numbers can be complex.

Let's say we have two funny looking 2x2 matrices. Just bear with me.

\begin{gather*}
	M_R=\begin{pmatrix}1 & 0 \\
	               0 & 1
	\end{pmatrix}\quad
	M_I=\begin{pmatrix}0 & -1 \\
	               1 & 0
	\end{pmatrix}
\end{gather*}

Now we can put two scalars \(a,b\) infront of these numbers and add it
up like we did before:

\begin{gather*}
	z=a\begin{pmatrix}1 & 0 \\
	               0 & 1
	\end{pmatrix}+b\begin{pmatrix}0 & -1 \\
	               1 & 0
	\end{pmatrix}\\
	\Rarr z=\begin{pmatrix}a & -b \\
	               b & a
	\end{pmatrix}
\end{gather*}

We can write this weird looking matrix differently:

\begin{gather*}
	z=a+bi
\end{gather*}

What is \(i\)? Well, we don't know. It's not a variable! But we know
what \(i^2\) is:

\begin{gather*}
	i^2=-1
\end{gather*}

Notice however much you can times \(a\), it won't affect \(b\), and vice
versa.

\subsection{Conjugate}\label{conjugate}
Say
\[
	z=a+bi
\]

Let the conjugate \(z^*\) of \(z\) be defined as this (we just
make the imaginary part negative)

\begin{gather*}
	z^*=a-bi=\begin{pmatrix}a  & +b \\
	               -b & a
	\end{pmatrix}
\end{gather*}

If you're writing in \hyperref[euler-form]{euler form}, then flip the sign of the angle
\[
	(e^{i\theta})^*=e^{-i\theta}
\]



\subsection{Euler Form}\label{euler-form}
Euler, another person, said that since \hyperref[taylor-series]{Taylor series} is a thing, we can say that

\begin{gather*}
	e^{i\theta}=1+i\theta+\frac{(i\theta)^2}{2!}+\frac{(i\theta)^3}{3!}+\cdots\\
	=1+i\theta-\frac{\theta^2}{2!}-i\frac{\theta^3}{3!}+\frac{\theta^4}{4!}+i\frac{\theta^5}{5!}
\end{gather*}

If we rearrange the terms we get

\begin{gather*}
	e^{i\theta}=\left(1-\frac{\theta^2}{2!}+\frac{\theta^4}{4!}-\cdots\right)+i\left(\theta -\frac{\theta^3}{3!}+\frac{\theta^5}{5!}+\cdots\right)\\
	=\cos\theta+i\sin\theta
\end{gather*}

This means we can represent \hyperref[complex-numbers]{complex numbers} in this weird form too:

\begin{gather*}
	z=re^{i\theta}
\end{gather*}

also trigometry can use complex numbers too
\[
	\cos\theta=
	\frac{1}{2}(e^{i\theta}+e^{-i\theta})
\]
\[
	\sin\theta=
	\frac{1}{2i}(e^{i\theta}-e^{-i\theta})
\]

\subsection{Complex Number Decomposition}\label{complex-number-decomposition}
This is all true
\[
	z = a+bi
\]
\[
	Re(z)=\frac{1}{2}(z+z^*)
\]
\[
	Im(z)=\frac{1}{2i}(z-z^*)
\]
\subsection{Absolute Value of Complex Number}\label{absolute-value-of-complex-number}
\[
	z\cdot z^*=|z|^2
\]
\subsection{Complex Number Trigonometry}\label{complex-number-trigonometry}
\[
	\cos\theta = \frac{e^{i\theta} + e^{-i\theta}}{2}
\]
\[
	\sin\theta = \frac{e^{i\theta} - e^{-i\theta}}{2i}
\]
\[
	\tan\theta = \frac{\sin\theta}{\cos\theta} = \frac{1}{i}\cdot\frac{e^{i\theta} - e^{-i\theta}}{e^{i\theta} + e^{-i\theta}} = -i\cdot\frac{e^{i\theta} - e^{-i\theta}}{e^{i\theta} + e^{-i\theta}}
\]


\section{Linear Algebra I}\label{linear-algebra-i}
More fun math with matrices
\subsection{Linear Independence}\label{linear-independence}

Let ``linearly independent states'' be states that you cannot create
another linearly independent state from multiplying and adding another
linearly independent state.

a cleaner definition is \(\ket{\psi_1},\ket{\psi_2},…\ket{\psi_n}\) are
linearly independent if the only way to get

\begin{gather*}
	c_1\ket{\psi_1}+c_2\ket{\psi_2}+\cdots+c_n\ket{\psi_n}=0
\end{gather*}

is having \(c_i=0\) for all (\(\forall\)) \(c_i\).

Example:

Given \(c\) is just a regular number,

\begin{gather*}
	\ket{+}=c\ket{-}\quad\text{will never happen how much you try}
\end{gather*}

and vice versa.

\subsection{Basis State}\label{basis-states}

If a set of states are BOTH

\begin{itemize}
	\tightlist
	\item
	      \hyperref[linear-independence]{linearly
		      independent}
	\item
	      span the whole space (i.e., every state in the space can be written as
	      a linear combination of these things)
\end{itemize}

then they are called basis states.

Note that \(\ket{+},\ket{-}\) are ``basis''
states

\subsection{Inner product I}\label{inner-product}

Inner product of states \(\ket{u}\) and \(\ket{v}\) is defined as

\begin{gather*}
	\ket{u}\cdot \ket{v}=u_1^*v_1+u_2^*v_2+...+u_n^*v_n
\end{gather*}

(if the entries are real numbers, the \hyperref[conjugate]{conjugates} do nothing and this
reduces to the usual dot product.)

\subsection{Orthogonality}\label{orthogonality}

State \(\ket{u},\ket{v}\) are orthogonal to each other if the
\hyperref[inner-product]{inner
	product} of state \(\ket{u}\) and \(\ket{v}\) are zero.

\begin{gather*}
	\ket{u}\cdot \ket{v}=0
\end{gather*}

They're also known as being perpendicular!


\subsection{Magnitude}\label{magnitude}

Where \(n\) is the dimensions, the magnitude of \(\ket{\psi}\) is
defined as:

\begin{gather*}
	||\ket{\psi}||=||\begin{pmatrix}\psi_1\\\psi_2\\\vdots\\\psi_n\end{pmatrix}||=\sqrt{|\psi_1|^2+|\psi_2|^2+...+|\psi_n|^2}
\end{gather*}

Note that
\[
	||\ket{\psi}||^2=\braket{\psi|\psi}
\]
You'll know what a bra-ket is later.

\subsection{Normal}\label{normal}

\(\ket{\psi}\) is normal if it has
\hyperref[magnitude]{magnitude}
1

\begin{gather*}
	||\ket{\psi}||=1
\end{gather*}

\subsection{Orthonormality}\label{orthonormality}

Let `` \(\ket{+},\ket{-}\) are orthonormal'' mean both
\(\ket{+},\ket{-}\) are both
\hyperref[orthogonality]{orthogonal}
and
\hyperref[normal]{normal}.

\subsection{Gram-Schmidt Procedure}\label{gram-schmidt-procedure}

If you have
\hyperref[linear-independence]{linearly
	independent} vectors and want to convert them into an
\hyperref[orthonormality]{orthonormal}
set, you can follow the Gram-Schmidt procedure.

Given \(v_1,v_2,…,v_n\) as
\hyperref[linear-independence]{LI}
\hyperref[vectors]{vectors},
then \(u_1, u_2, …, u_n\) can be built using

\begin{gather*}
	u_k=v_k-\sum_{i=1}^{k-1}\Pi_{u_i}(v_k)
\end{gather*}

This makes \(u\) an
\hyperref[orthogonality]{orthogonal}
\hyperref[vectors]{vector}

Then you
\href{https://www.notion.so/Going-from-1-1-2-to-Understanding-Quantum-Mechanics-344d7c8acbf380088bbdfa6e9ac17c16?pvs=21}{normalize}
\(u\) to get a
\hyperref[orthonormality]{orthonormal}
\hyperref[vectors]{vector}.
You do this by multiplying the
\hyperref[magnitude]{magnitude}
of the
\hyperref[vectors]{vector}.
\subsection{Bra, Transpose}\label{bratranspose}

We just learned what ket \(\ket{x}\) means --- a \(N\times 1\) vector.
Let that ALSO be called a ket. If we flip it horizontally, we get a bra
\(\bra{x}\).

If a ket is

\begin{gather*}
	\ket{x}=\begin{pmatrix}
		1+3i \\
		2+5i \\
		7
	\end{pmatrix}\quad \ket{\psi}=\begin{pmatrix}
		9 \\
		9 \\
		0 \\
		0 \\
		0
	\end{pmatrix}\quad\ket{\phi}=\begin{pmatrix}
		0 \\
		1
	\end{pmatrix}
\end{gather*}

Then it's counterpart is a bra

\begin{gather*}
	\bra{x}=\begin{pmatrix}
		1-3i &
		2-5i &
		7
	\end{pmatrix}\quad \bra{\psi}=\begin{pmatrix}
		9 &
		9 &
		0 &
		0 &
		0
	\end{pmatrix}\quad\bra{\phi}=\begin{pmatrix}
		0 &
		1
	\end{pmatrix}
\end{gather*}

You swap the columns for the rows (transpose!) and make every \hyperref[element]{element}
the \hyperref[conjugate]{conjugate} of itself. It becomes a \(1\times N\) matrix instead.

Let this weird transformation be known as a conjugate transpose, denoted
as

\begin{gather*}
	\bra{v}=(\ket{v})^+
\end{gather*}


\subsection{Kronecker Delta}\label{kronecker-delta}

Where \(\delta_{ij}\) is just something we define as
\[\delta_{ij}=\begin{cases}1\quad i=j\\0\quad i\neq j\end{cases}\]

\subsection{Bra-ket, matrix
	multiplication}\label{bra-ket-matrix-multiplication}

Now let this be a bra-ket (get it?)

\begin{gather*}
	\braket{\phi|\psi}=\bra{\phi}\ket{\psi}=\begin{bmatrix}c^* & d^*\end{bmatrix}\begin{bmatrix}a\\b\end{bmatrix}\quad
\end{gather*}

This means we're multiplying these two matrices
\(\begin{bmatrix}c^* & d^*\end{bmatrix},\begin{bmatrix}a\\b\end{bmatrix}\)
together.

The way we multiply matrices is by performing this algorithm.

Let there be two matrices, \(A,B\) with size \(m\times n\) and
\(n\times p\) respectively. We get a matrix \(C=AB\) which is size
\(m\times p\).

for an \hyperref[element]{element} on row \(i\), column \(j\) in \(C\) we let this be called
\(C_{ij}\)

\begin{gather*}
	C_{ij}=\sum_{k=1}^nA_{ik}B_{kj}
\end{gather*}

so

\begin{gather*}
	\braket{\phi|\psi}\\=\bra{\phi}\ket{\psi}\\=\begin{pmatrix}c^* & d^*\end{pmatrix}\begin{pmatrix}a\\b\end{pmatrix}\\=\begin{pmatrix}
		c^*a+d^*b\end{pmatrix}\\=c^*a+d^*b
\end{gather*}

It is a 1x1 matrix which can be simplified to a scalar value.

Note that

\begin{gather*}
	\braket{x+|x+}=\braket{x-|x-}=1\\
	\braket{x+|x-}=\braket{x-|x+}=0
\end{gather*}

Where \(x+,x-\) are \hyperref[orthonormality]{orthonormal} to each other. Another way we can say
this is \(\{\ket{x+},\ket{x-}\}\) forms an \hyperref[orthonormality]{orthonormal} \hyperref[basis-states]{basis}.

Let this be the notation of if/else

\begin{gather*}
	x=\begin{cases}1\quad\text{if true}\\0\quad\text{if false}\end{cases}
\end{gather*}

a set of \hyperref[vectors]{vectors} \(\{v_1,v_2,...,v_n\}\) form an \hyperref[orthonormality]{orthonormal} \hyperref[basis-states]{basis} if
and only if

\begin{gather*}
	\braket{v_i|v_j}=\delta_{ij}=\begin{cases}1\quad i=j\\0\quad i\neq j\end{cases}
\end{gather*}

\subsection{Square Matrix}\label{square-matrix}

a square matrix is a matrix with the same number of rows and columns
(i.e., an \(N\times N\) matrix)

\subsection{Symmetric Matrix}\label{symmetric-matrix}

A matrix \(X\) where

\begin{gather*}
	X=X^T
\end{gather*}

(note that \(X^T\) means
\href{https://www.notion.so/Going-from-1-1-2-to-Understanding-Quantum-Mechanics-344d7c8acbf380088bbdfa6e9ac17c16?pvs=21}{transpose})



\subsection{Vector}\label{vectors}

Let vector just be the same thing as
\hyperref[ket-states]{states}.

\subsection{Vector Space}\label{vector-spaces}

Let a vector space \(V\) be a set of \hyperref[vectors]{vectors} with

\begin{itemize}
	\tightlist
	\item
	      addition such that
\end{itemize}

\begin{gather*}
	\ket{u}+\ket{v}\in V\quad ;\quad\ket{u},\ket{v}\in V
\end{gather*}

\begin{itemize}
	\tightlist
	\item
	      scalar multiplication such that
\end{itemize}
\subsection{Operator}\label{operators}

Let \(\hat{T}\) be an operator.

An operator is a
\hyperref[functions]{function}
where

\begin{gather*}
	\hat{T}:V\rarr V
\end{gather*}

where V are the set of all possible \hyperref[vectors]{vectors} in a \hyperref[hilbert-spaces]{Hilbert space}.

In quantum mechanics, all observed operators are linear which means

\begin{gather*}
	\hat{T}(a\ket{u}+b\ket{v})=a\hat{T}(\ket{u})+b\hat{T}(\ket{v})
\end{gather*}

If we were to observe a non-linear operator, then that would cause wacky things.

also they can be represented as \hyperref[matrices]{matrices}

\begin{gather*}
	\hat{T}\ket{\psi}=\sum_{m}\psi_m'\ket{m}
\end{gather*}

where

\begin{gather*}
	\begin{bmatrix}
		\psi_1' \\
		\psi_2' \\
		\vdots  \\
		\psi_d'
	\end{bmatrix}=
	\begin{bmatrix}
		T_{11} & T_{12} & \cdots & T_{1d} \\
		T_{21} & T_{22} & \cdots & T_{2d} \\
		\vdots                            \\
		T_{d1} & T_{d2} & \cdots & T_{dd}
	\end{bmatrix}\begin{bmatrix}\psi_1 \\ \psi_2 \\ \vdots \\\psi_d\end{bmatrix}
\end{gather*}

where \(\ket{m}\) is the basis for \(\ket{\psi}\)

Operators must be square.

\begin{gather*}
	a\ket{v}\in V\quad ;\quad a\in \mathbb{F},\ket{v}\in V
\end{gather*}

\section{Hilbert Space}\label{hilbert-spaces}

A Hilbert space \(H\) describes quantum space.

\begin{gather*}
	H
\end{gather*}

A Hilbert space is a \hyperref[vector-spaces]{vector space}, possibly infinite-dimensional.

\(H\) has an \hyperref[inner-product]{inner product} and is complete.

The states we wrote before for example \(\ket{\psi}\) are \hyperref[states]{states} which
exist in a Hilbert space.

\subsection{Inner Product II}\label{inner-product-ii}

building on top of what we said about the \hyperref[inner-product]{inner product},

an inner product is any

\begin{gather*}
	\langle u|v\rangle
\end{gather*}

which is

\begin{enumerate}
	\def\labelenumi{\arabic{enumi}.}
	\tightlist
	\item
	      \hyperref[conjugate]{conjugate} symmetric
\end{enumerate}

\begin{gather*}
	\braket{u|v}^*=\braket{v|u}
\end{gather*}

Example

let

\begin{gather*}
	\ket{\psi}=\begin{bmatrix}a\\b\end{bmatrix}\quad \Rarr \bra{\psi}=\begin{bmatrix}a^* & b^*\end{bmatrix}
	\\
	\ket{\phi}=\begin{bmatrix}c\\d\end{bmatrix}\Rarr \bra{\phi}=\begin{bmatrix}c^* & d^*\end{bmatrix}
\end{gather*}

so

\begin{gather*}
	(\braket{\psi|\phi})^*=\left(\begin{bmatrix}a^* & b^*\end{bmatrix}\begin{bmatrix}c\\d\end{bmatrix}\right)^*=(a^*c+b^*d)^*=ac^*+bd^*=c^*a+d^*b\\
	=\braket{\phi|\psi}
\end{gather*}

so \(\phi,\psi\) are conjugate symmetric

\begin{enumerate}
	\def\labelenumi{\arabic{enumi}.}
	\tightlist
	\item
	      linear
\end{enumerate}

\begin{gather*}
	\langle u|(a|v\rangle + b|w\rangle) = a\langle u|v\rangle + b\langle u|w\rangle
\end{gather*}

\begin{enumerate}
	\def\labelenumi{\arabic{enumi}.}
	\tightlist
	\item
	      positive definite
\end{enumerate}

\begin{gather*}
	\forall |u\rangle \neq 0, \quad \langle u|u\rangle > 0
\end{gather*}

\subsection{Completeness}\label{completeness}

Don't worry about it. -prof. Zhou
\section{Planck-Einstein Relation}\label{planck-einstein-relation}

Planck, a person, first proposed (and Einstein, another person, later
applied to photons via the photoelectric effect) that

\begin{gather*}
	E=hf
\end{gather*}

It means energy \(E\) of a photon is Planck's constant
\(h=6.626\times 10^{-34}Js\) times the frequency of the light \(f\).

They are all \hyperref[variable]{variables} so we can substitute them with what we want.
Because it was experimental, we can't really prove with math that this
is true, but we proved it with experiments. This is called the
Planck--Einstein relation.

\section{Angular Frequency, reduced Planck
  constant}\label{angular-frequency-reduced-planck-constant}

Some dude in physics was like let's make a new, transformed version of
frequency called angular frequency \(\omega\)! (Not to be confused with
angular momentum, which is a different physical quantity.)

Let

\begin{gather*}
	\omega=2\pi f
\end{gather*}

Because of this, we can say that

\begin{gather*}
	E=hf\\
	=h\frac{\omega}{2\pi}
\end{gather*}

Let the reduced Planck constant be

\begin{gather*}
	\hbar=\frac{h}{2\pi}
\end{gather*}

so we can also say

\begin{gather*}
	E=\hbar \omega
\end{gather*}

Bohr, a person, hypothesized that atoms absorb light at discrete
frequencies.

This means that atoms (photons too) have discrete (non-continuous)
energy levels.

\section{Particle Energy}\label{particle-energies}

We just said photons have discrete energy levels. Let's say they have
\(k\) energy levels.

We know experimentally, that any state \(\ket{\psi(t)}\) (e.g., spin) of
a quantum particle can be expanded in the energy basis. Let the energy
basis be just a set of \hyperref[basis-states]{basis states} \(E_1,E_2, …, E_k\). This means a
photon can be at those energies.

A state is a superposition of all those energies. Recall that a super
position is just a sum (\(\sum\)) of those energies. I put \(k\) under
the sum showing we must add all the energies and their corresponding
\(c_k\) values (aka. coefficients) together.

\begin{gather*}
	\ket{\psi(0)}=\sum_kc_k\ket{E_k}
\end{gather*}

\section{Particle Phase}\label{particle-phases}

We also just proved that \(E=\hbar\omega\) so that particle with energy
\(\ket{E_k}\) will oscillate at frequency

\begin{gather*}
	\omega_k=\frac{E_k}{\hbar}
\end{gather*}

Oscillating means the phase of the particle changes so we can represent
the phase as a \hyperref[complex-numbers]{complex number}.

Let's say there is a two-level \hyperref[quantum-system]{quantum system} in some superposition.
This just means

\begin{gather*}
	\ket{\psi}=\alpha\ket{0}+\beta\ket{1}=\begin{pmatrix} \alpha\\\beta\end{pmatrix}
\end{gather*}

Because \(\alpha,\beta\) can be complex, we proved experimentally that a
glass polarizer, like a polarized sunglass, can change the phase of the
state \(\ket{1}\) for example.

Let

\begin{gather*}
	\beta'=\beta e^{i\delta}
\end{gather*}

Where \(\delta\) can be any phase angle from 0 to \(2\pi\).

This is the same phase as which destructively interfere in the double
slit experiment. That's why phases exist. Because light is a wave with a
frequency (and also a particle!)


\section{Schrödinger's Equation, Hamiltonian I,
 }\label{schruxf6dingers-equation-hamiltonian}

Recall from \hyperref[particle-energies]{particle energies} that

\begin{gather*}
	\ket{\psi(0)}=\sum_kc_k\ket{E_k}
\end{gather*}

Because we learned this thing called \hyperref[particle-phases]{particle phases}, we can say that particle
oscillate at different phases. This is shown using phases

\begin{gather*}
	\ket{\psi(t)}=\sum_kc_ke^{-i\omega_kt}\ket{E_k}\\
	=\sum_kc_ke^{-iE_kt/\hbar}\ket{E_k}\\
\end{gather*}

If we take the derivative from both sides

\begin{gather*}
	\frac{d}{dt}\ket{\psi(t)}=\frac{d}{dt}\left[\sum_kc_ke^{-iE_kt/\hbar}\ket{E_k}\right]\\
	=\sum_kc_k\cdot \left(-i\frac{E_k}{\hbar}\right)e^{-iE_kt/\hbar}\ket{E_k}\\
	=-\frac{i}{\hbar}\sum_kc_k\cdot E_ke^{-iE_kt/\hbar}\ket{E_k}
\end{gather*}

Let the Hamiltonian \(H\)be an \hyperref[operators]{operator} where this is true

\begin{gather*}
	H\ket{E_k}=E_k\ket{E_k}
\end{gather*}

so

\begin{gather*}
	=-H\frac{i}{\hbar}\sum_kc_k\cdot e^{-iE_kt/\hbar}\ket{E_k}\\
	=-\frac{i}{\hbar}H\ket{\psi(t)}
\end{gather*}

which gives us Schrödinger's equation (in state-vector form)

\begin{gather*}
	i\hbar\frac{d}{dt}\ket{\psi(t)}=H\ket{\psi(t)}
\end{gather*}




\section{Quantum System}\label{quantum-system}
A quantum system is defined by:
\begin{enumerate}
	\def\labelenumi{\arabic{enumi}.}
	\tightlist
	\item A \hyperref[hilbert-spaces]{Hilbert space} \(H\)
	\item A Hamiltonian \(\hat{H}\) generating time evolution via the \hyperref[schruxf6dingers-equation-hamiltonian]{Schrödinger equation}.
\end{enumerate}
The Hilbert space's dimensions defines the fixes the spin of the system.

\subsection{Spin-1/2 System}\label{spin-12-systems}
This spin is for 2-dimensional \hyperref[hilbert-spaces]{Hilbert spaces}.
Electrons, protons, neutrons and quarks have spin-1/2.
For spin-1/2 systems,
\[
	S_x\triangleq\frac{\hbar}{2}\sigma_x\quad
	S_y\triangleq\frac{\hbar}{2}\sigma_y\quad
	S_z\triangleq\frac{\hbar}{2}\sigma_z
\]
i.e.,
\[
	S_x\triangleq\frac{\hbar}{2}\begin{pmatrix}0 & 1\\ 1 & 0\end{pmatrix}\quad
	S_y\triangleq\frac{\hbar}{2}\begin{pmatrix}0 & -i\\ i & 0\end{pmatrix}\quad
	S_z\triangleq\frac{\hbar}{2}\begin{pmatrix}1 & 0\\ 0 & -1\end{pmatrix}
\]
Note that
\[
	\ket{z+}=\ket{\uparrow} = \ket{0}=\begin{pmatrix}1 \\ 0\end{pmatrix}\quad
	\ket{z-}=\ket{\downarrow} = \ket{1}=\begin{pmatrix}0 \\ 1\end{pmatrix}
\]
And
\[
	\ket{+}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})=\frac{1}{\sqrt{2}}\begin{pmatrix}1 \\ 1\end{pmatrix}\quad
	\ket{-}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})=\frac{1}{\sqrt{2}}\begin{pmatrix}1 \\ -1\end{pmatrix}
\]

\subsection{Spin-1 System}\label{spin-1-systems}
This spin is for 3-dimensional \hyperref[hilbert-spaces]{Hilbert spaces}.
W/Z bosons, photons and gluons have spin-1.
For spin-1 systems,
\[
	S_x\triangleq\frac{\hbar}{2}\Sigma_x\quad
	S_y\triangleq\frac{\hbar}{2}\Sigma_y\quad
	S_z\triangleq\frac{\hbar}{2}\Sigma_z
\]
i.e.,
\[
	S_x\triangleq\frac{\hbar}{\sqrt{2}}\begin{pmatrix}0 & 1 & 0\\ 1 & 0 & 1\\ 0 & 1 & 0\end{pmatrix}\quad
	S_y\triangleq\frac{\hbar}{\sqrt{2}}\begin{pmatrix}0 & -i & 0\\ i & 0 & -i\\ 0 & i & 0\end{pmatrix}\quad
	S_z\triangleq\hbar\begin{pmatrix}1 & 0 & 0\\ 0 & 0 & 0\\ 0 & 0 & -1\end{pmatrix}
\]
Note that
\[
	\ket{1}=\begin{pmatrix}1 \\ 0 \\ 0\end{pmatrix}\quad
	\ket{0}=\begin{pmatrix}0 \\ 1 \\ 0\end{pmatrix}\quad
	\ket{-1}=\begin{pmatrix}0 \\ 0 \\ -1\end{pmatrix}\quad
\]
For photons, $S_z$ is often called rectilinear and $S_x$ is often called diagonal and $S_y$ are often called circular.



\section{Linear Algebra II}\label{linear-algebra-ii}



\subsection{Trace}\label{traces}

Tr(A) is only defined for a \hyperref[square-matrix]{square matrix}.

Given a square matrix \(A\),

\begin{gather*}
	tr(A)=a_{11}+a_{22}+...+a_{nn}
\end{gather*}

Note that

\begin{gather*}
	tr(AB)=tr(BA)
\end{gather*}

Traces are just adding diagonals together

\begin{gather*}
	tr(\hat{T})=\sum_n\bra{n}\hat{T}\ket{n}=\sum_nT_{nn}
\end{gather*}

Properties:

\begin{gather*}
	tr(\hat{A}\hat{B})=tr(\hat{B}\hat{A})\\
	tr(\hat{A}\hat{B}\hat{C})=tr(\hat{B}\hat{C}\hat{A})=tr(\hat{C}\hat{A}\hat{B})
\end{gather*}

Note that \(tr(\hat{A}\hat{B}\hat{C})\neq tr(\hat{A}\hat{C}\hat{B})\) →
it must be cyclical permutation

\begin{gather*}
	tr(\alpha\hat{A}+\beta\hat{B})=\alpha tr(\hat{A})+\beta Tr(\hat{B})
\end{gather*}

Also note that if you get the traces of a \(1\times 1\) matrix, it is just

\[
	tr(\begin{pmatrix}a\end{pmatrix})=a
\]

\subsection{Determinant}\label{determinants}

Det is a function that we can apply to matrices to get a scalar. It's
defined for \(n\times n\) only.

For a \(2\times 2\) matrix it's just

\begin{gather*}
	det\begin{bmatrix}a & b \\ c & d\end{bmatrix}=ad-bc
\end{gather*}

For an \(n\times n\) matrix it can be done with cofactor expansion

\begin{gather*}
	det(A)=\sum_{j=1}^n(-1)^{i+j}a_{ij}M_{ij}
\end{gather*}

Note it's recursive so it's really expensive

also dont forget that

\begin{gather*}
	det(\hat{A}\hat{B})=det(\hat{A})det(\hat B)
\end{gather*}

An easier way to do it is through row reduction. Make a triangular
matrix and multiply the diags together

\begin{gather*}
	det\begin{pmatrix}a_{11} & *      & *      \\
	                      & a_{22} & *      \\
	               0      & 0      & a_{33}
	\end{pmatrix}=a_{11}a_{22}a_{33}
\end{gather*}

\subsection{Matrix
	Invertibility}\label{invertibility-of-matrices}

\hyperref[operators]{Operator} \(\hat{T}\) (i.e., a matrix) is invertible if and only if
\(det(\hat{T})\neq0\)

\subsection{Eigenvector/Eigenvalue}\label{eigenvectorseigenvalues}

\url{https://www.youtube.com/watch?v=PFDu9oVAE-g}

\begin{gather*}
	A\vec{v}=\lambda\vec{v}
\end{gather*}

Basically if I apply matrix transformation \(A\) to a vector \(\ket{v}\)
and it still points in the same direction, it is an Eigenvector. It
could be multiplied by a scalar value \(\lambda\) (Eigenvalue) after
applying the transform, which means its \hyperref[magnitude]{magnitude} need not be preserved
by the transform.

The set of all vector \(\ket{v}\)'s are Eigenvectors of transformation
\(A\).

An \hyperref[eigenstates]{Eigenstate} are states instead of \hyperref[vectors]{vectors}, \hyperref[operators]{operators} instead of
matrices but are the same thing.

\subsection{Eigenstates}\label{eigenstates}
Eigenstates are \hyperref[eigenvectorseigenvalues]{eigenvectors} of \hyperref[operators]{operators}.

\begin{gather*}
	\hat{A}\ket{\psi}=\lambda\ket{\psi}
\end{gather*}

\subsection{Eigenspace}\label{eigenspaces}
An eigenspace of operator \(\hat{A}\) is the set of all \hyperref[eigenvectorseigenvalues]{eigenvectors}/\hyperref[eigenstates]{eigenstates} with the same \hyperref[eigenvectorseigenvalues]{eigenvalue} plus the zero vector.

Example

If \hyperref[operators]{operator} \(\hat{A}\) has \hyperref[eigenvectorseigenvalues]{eigenvalue} \(\lambda\), then the \(\lambda\)-eigenspace of the operator \(\hat{A}\) is

\[
	\{\ket{\psi}:\hat{A}\ket{\psi}=\lambda\ket{\psi}\}
\]

\subsection{Eigenbasis}\label{eigenbasis}
An eigenbasis of \hyperref[operators]{operator} \(\hat{A}\) is a set of \hyperref[eigenvectorseigenvalues]{eigenvectors}/\hyperref[eigenstates]{eigenstates} that form a \hyperref[basis-states]{basis} for the \hyperref[hilbert-spaces]{Hilbert space}.

Every \hyperref[hermitian-operators]{Hermitian operator} has an \hyperref[eigenbasis]{eigenbasis}.

\subsection{Characteristic Polynomial}\label{characteristic-polynomial}

polynomial
\(p(\lambda)=\alpha_0\lambda^n+\alpha_1\lambda^{n-1}+...+\alpha_{n-1}\lambda^{1}+\alpha_n\)

\begin{gather*}
	p(\lambda)=det(A-\lambda I)
\end{gather*}

Solving

\begin{gather*}
	p(\lambda)=0
\end{gather*}

Gives \(\lambda\) values which are \hyperref[eigenvectorseigenvalues]{eigenvalues} of A

\subsection{Subspace}\label{subspaces}

A
\hyperref[subsets]{subset}
\(U\subseteq V\) is a subspace if it satisfies:

\begin{enumerate}
	\def\labelenumi{\arabic{enumi}.}
	\tightlist
	\item
	      contains zero vector
	\item
	      closed under addition \(\ket{u},\ket{v}\in U, \ket{u}+\ket{v}\in U\)
	\item
	      closed under scalar multiplication
	      \(\ket{u}\in U, c\in \mathbb{F}\Rarr c\ket{u}\in U\)
\end{enumerate}

\subsection{Span Function}\label{span-function}

A span of \hyperref[vectors]{vectors} is the set of all linear combinations you can build
from them


\subsection{Invariant Subspace}\label{invariant-subspaces}

A
\hyperref[subspaces]{subspace}
\(U\subseteq V\) is called \(\hat{T}\)-invariant if \hyperref[operators]{operator} \(\hat{T}\)
keeps everything in U inside U. If it is T-invariant, then this must
always be true:

\begin{gather*}
	\hat{T}(U)=\{\hat{T}\ket{\psi}|\ket{\psi}\in U\}\subseteq U
\end{gather*}

Simplest T invariant subspace is (with dimensions \(d\))

\begin{gather*}
	U=span\{\ket{u}\}\\
	\hat{T}\ket{y}=\lambda \ket{u}
\end{gather*}

When \(\lambda\) is \hyperref[eigenvectorseigenvalues]{Eigenvalue} and \(\ket{u}\) is \hyperref[eigenvectorseigenvalues]{eigenvector} and
\(\ket{u}\neq 0\) and you can define these \hyperref[vectors]{vectors} by checking

\begin{gather*}
	(\hat{T}-\lambda I)\ket{u}=0\\
	0=det(\hat{T}-\lambda I)\\
	=(\lambda_1-\lambda)(\lambda_2-\lambda)..(\lambda_d-\lambda)
\end{gather*}

In general, this \hyperref[determinants]{determinant} has degree \(\lambda_i\) are zeroes of
character polynomial \(\lambda_i\in C\) can be repeated

\subsection{Diagonalization}\label{diagonalization}

A matrix is diagonalizable if it has \hyperref[linear-independence]{linearly independent} \hyperref[eigenvectorseigenvalues]{eigenvectors}
which occurs if the geometric multiplicity (dimension of the \hyperref[eigenspaces]{eigenspace})
equals the algebraic multiplicity (root multiplicity in the
\hyperref[characteristic-polynomial]{characterstic polynomial}) for every \hyperref[eigenvectorseigenvalues]{eigenvalue}.

If matrix \(A\) is diagonalizable then it can be written as

\begin{gather*}
	A=PDP^{-1}
\end{gather*}

Where \(D=diag(\lambda_1, … , \lambda_n)\)

\subsection{Matrix Exponential}\label{matrix-exponentials}

The matrix exponential of
\hyperref[square-matrix]{square
	matrix} \(X\) is defined by a power series

\begin{gather*}
	e^X=\sum_{n=0}^\infin\frac{X^n}{n!}
\end{gather*}

Properties of this

\begin{itemize}
	\tightlist
	\item
	      differential equations

	      \begin{itemize}
		      \tightlist
		      \item
		            the function \(U(t)=e^{tX}\) is the only unique soln. to
	      \end{itemize}
\end{itemize}

\begin{gather*}
	\frac{dU}{dt}=XU(t)\quad U(0)=I
\end{gather*}

\begin{itemize}
	\tightlist
	\item
	      diagonalizable matrices → if \(X\) is diagonalizable (which is
	      guaranteed for, e.g., real \hyperref[symmetric-matrix]{symmetric} or \hyperref[hermitian-matrix]{Hermitian matrices}) then
	      we can write \(X=PDP^{-1}\) with
	      \(D=diag(\lambda_1, … \lambda_n)\) then
\end{itemize}

\begin{gather*}
	e^X=Pdiag(e^{\lambda_1},...,e^{\lambda_n})P^{-1}
\end{gather*}

where \(P\) is the basis matrix --- a matrix whose columns are the
\hyperref[eigenvectorseigenvalues]{eigenvectors} \(X\) w.r.t \(\hat{A}\).

Specifically, if \(v_1,v_2,…,v_n\) are \hyperref[eigenvectorseigenvalues]{eigenvectors} of \(X\)
corresponding to \hyperref[eigenvectorseigenvalues]{eigenvalues} \(\lambda_1,\lambda_2,…,\lambda_n\) then

\begin{gather*}
	P = \begin{bmatrix} | & | & & | \\ v_1 & v_2 & \cdots & v_n \\ | & | & & | \end{bmatrix}
\end{gather*}

\begin{itemize}
	\tightlist
	\item
	      \hyperref[hermitian-matrix]{Hermitian} and skew-Hermitian matrices → if \(H=H^+\) then \(e^{iH}\)
	      is \hyperref[unitary-operator]{unitary} this means
\end{itemize}

\begin{gather*}
	(e^{iH})^+e^{iH}=I
\end{gather*}

Note that if $A$ is a matrix, then, via \hyperref[taylor-series]{Taylor series}, we have
\[e^{At} = \sum_{k=0}^{\infty}\frac{A^k t^k}{k!} = I + At + \frac{(At)^2}{2!} + \frac{(At)^3}{3!} + \cdots\]

\[\frac{d}{dt}e^{At} = \sum_{k=1}^{\infty}\frac{kA^k t^{k-1}}{k!} = A\sum_{k=1}^{\infty}\frac{A^{k-1}t^{k-1}}{(k-1)!} = Ae^{At} = e^{At}A\]

\subsection{Commutator}\label{commutators}

The commutator of two \hyperref[operators]{operators} \(\hat{A},\hat{B}\) is defined as

\begin{gather*}
	[\hat{A},\hat{B}]=\hat{A}\hat{B}-\hat{B}\hat{A}
\end{gather*}

If \([\hat{A},\hat{B}]=0\) then that means you can commute
\hyperref[bra-ket-matrix-multiplication]{matrix
	multiplication}.

If "\(\hat{A}\) commutes with \(\hat{B}\)", then that means \(\hat{A}\hat{B}=\hat{B}\hat{A}\) which also means \([\hat{A},\hat{B}]=0\).

\subsection{Adjoint}\label{adjoint}

An adjoint is basically a \hyperref[conjugate]{conjugate} transpose but for an \hyperref[operators]{operator}
(whereas a conjugate transpose can only be applied to a matrix)

Given operator \(\hat{A}\), it's adjoint is the operator \(\hat{A}^+\)
such that \(\forall\ket{i},\ket{j}\)

\begin{gather*}
	\bra{i}\hat{A}^+\ket{j}=(\bra{j}\hat{A}\ket{i})^*\\
	(\hat{A}^+)_{ij}=\hat{A}^*_{ji}
\end{gather*}

Where, for any operator \(\hat{O}\), a matrix element
\(\bra{i}\hat{O}\ket{j}\) is just the \((i,j)\)-th element of the matrix
representing \(\hat{O}\) in basis \(\{\ket{i}\}\)

Note that \(\ket{i}\) and \(\ket{j}\) are arbitrary \hyperref[basis-states]{basis states}

where \(A^*\) is the complex conjugation of \(A\)

properties:

\begin{gather*}
	(\hat{A}^+)^+=\hat{A}\\
	(\hat{A}+\hat{B})^+=\hat{A}^++\hat{B}^+\\
	(\hat{A}\hat{B})^+=\hat{B}^+\hat{A}^+
\end{gather*}

Example:

Consider operator

\begin{gather*}
	\hat{A}=\ket{1}\bra{2}+2i\ket{2}\bra{1}+3\ket{2}\bra{2}\\
	\Rarr \text{adjoint of $\hat{A}$}=\ket{2}\bra{1}-2i\ket{1}\bra{2}+3\ket{2}\bra{2}
\end{gather*}

So w.r.t. to basis \(\{\ket{1},\ket{2}\}\), \(\hat{A}\) is

\begin{gather*}
	\hat{A}=\begin{pmatrix}
		0  & 1 \\
		2i & 3
	\end{pmatrix}
\end{gather*}

The conjugate transpose of \(\hat{A}=\hat{A}^+\) which is

\begin{gather*}
	\hat{A}^+=\begin{pmatrix}
		0 & -2i \\
		1 & 3
	\end{pmatrix}
\end{gather*}

Which, once we pick an orthonormal basis, the matrix of \(\hat{A}^+\)
equals the conjugate transpose of the matrix of \(\hat{A}\)

\begin{gather*}
	\hat{A}^+=\text{adjoint of $\hat{A}$}
\end{gather*}

\subsection{Normal Operator}\label{normal-operator}

An \hyperref[operators]{operator} $A$ is normal if and only if

\begin{gather*}
	\hat{A}\hat{A}^+=\hat{A}^+\hat{A}
\end{gather*}

Not to be confused with \hyperref[normal]{normal}.

\subsection{Hermitian Matrix}\label{hermitian-matrix}

Let a ``Hermitian matrix'' be a matrix that equals its own \hyperref[conjugate]{conjugate}
transpose

If and only if \(A\) is a matrix and is Hermitian, then this is true

\begin{gather*}
	A=A^+
\end{gather*}

This means a bra and ket are Hermitian matrices of each other!

Another example:

\begin{gather*}
	\ket{\psi}=\begin{bmatrix}a\\b\end{bmatrix}\quad \ket{\phi}=\begin{bmatrix}c\\d\end{bmatrix}\Rarr \bra{\phi}=\begin{bmatrix}c^* & d^*\end{bmatrix}
\end{gather*}

\subsection{Hermitian Operator}\label{hermitian-operators}

An \hyperref[operators]{operator} is \hyperref[hermitian-matrix]{Hermitian} if and only if the \hyperref[adjoint]{adjoint} of the operator is
equal to itself.


Properties

\begin{itemize}
	\tightlist
	\item
	      all \hyperref[eigenvectorseigenvalues]{eigenvalues} of a Hermitian operator are real
	\item
	      All \hyperref[eigenvectorseigenvalues]{eigenvectors} of a Hermitian operator form an \hyperref[orthonormality]{orthonormal} \hyperref[basis-states]{basis}
	\item
	      \hyperref[observables]{observables} are usually some kind of Hermitian operator
\end{itemize}

The \hyperref[eigenvectorseigenvalues]{eigenvalues} represent possible measurement outcomes. \hyperref[orthonormality]{Orthonormal}
\hyperref[eigenstates]{eigenstates} \(\ket{\alpha}\) mean that these are definite value states


\subsection{Anti-Hermitian Operator}\label{anti-hermitian-operators}

An \hyperref[operators]{operator} is anti-hermitian if and only if

\begin{gather*}
	\hat{B}^+=-\hat{B}
\end{gather*}

note that if you do

\begin{gather*}
	\hat{A}=i\hat{B}
\end{gather*}

then \(\hat{A}\) is \hyperref[hermitian-operators]{Hermitian}

Any operator \(\hat{T}\) can be written as linear combination of
\hyperref[hermitian-operators]{Hermitian} and anti-Hermitian operator.


\begin{gather*}
	\hat{T}=\hat{A}+i\hat{B}\\
	\hat{A}=\frac{\hat{T}+\hat{T}^+}{2}\\
	\hat{B}=\frac{\hat{T}-\hat{T}^+}{2i}
\end{gather*}

Note that this is true for \hyperref[complex-number-decomposition]{regular complex numbers}
\subsection{Positive Semi-Definite (PSD)
	Operator}\label{positive-semi-definite-psd-operators}

An operator \(\hat{P}\) is PSD if and only if

\begin{itemize}
	\tightlist
	\item
	      \(\hat{P}\) is \hyperref[hermitian-operators]{Hermitian}
	\item
	      this holds true for all states \(\ket{\psi}\) :
\end{itemize}

\begin{gather*}
	\bra{\psi}\hat{P}\ket{\psi}\geq 0
\end{gather*}

This means the angle between \(\ket{\psi}\) and \(\hat{P}\ket{\psi}\) is
less than 90 degrees.

Example

\begin{itemize}
	\tightlist
	\item
	      \hyperref[projectors]{Projectors} are PSD
\end{itemize}

\subsection{Projection}\label{projections}

The projection of \(v\) onto a
\hyperref[vectors]{vector}
\(u\) is

\begin{gather*}
	\Pi_u(v)=\frac{\braket{v|u}}{\braket{u|u}}u
\end{gather*}

Example

\begin{gather*}
	\ket{\psi}=\begin{bmatrix}a\\b\end{bmatrix}\quad \Rarr \bra{\psi}=\begin{bmatrix}a^* & b^*\end{bmatrix}
	\\
	\ket{\phi}=\begin{bmatrix}c\\d\end{bmatrix}\Rarr \bra{\phi}=\begin{bmatrix}c^* & d^*\end{bmatrix}
\end{gather*}

\begin{gather*}
	\Pi_\psi(\phi)=\frac{\begin{bmatrix}c^* & d^*\end{bmatrix}\begin{bmatrix}a\\b\end{bmatrix}}{\begin{bmatrix}a^* & b^*\end{bmatrix}\begin{bmatrix}a\\b\end{bmatrix}}\begin{bmatrix}a\\b\end{bmatrix}\\
	=\frac{c^*a+d^*b}{a^*a+b^*b}\begin{bmatrix}a\\b\end{bmatrix}
\end{gather*}

This is similar to the dot product, but it gives a
\hyperref[vectors]{vector}

\subsection{Projector}\label{projectors}
Projectors are
\hyperref[operators]{operators}
with a special property

\begin{gather*}
	\hat{\Pi}^2=\hat{\Pi}
\end{gather*}

A projector operator (aka. a projector) is a
\hyperref[projections]{projection}
that

\begin{gather*}
	\hat{\Pi}_u=\frac{\ket{u}\bra{u}}{\braket{u|u}}
\end{gather*}

Usually, by convention, we pick the representative state \(\ket{u}\) to be normalized, i.e., \(\braket{u|u}=1\).
So sometimes it can be written as
\[
	\hat{\Pi}_u=\ket{u}\bra{u}\quad \text{if} \quad\braket{u|u}=1
\]

once i apply it to a state \(v\), i get the projection of that state
\(v\) on state \(u\)

\subsection{Unitary Operator}\label{unitary-operator}

An \hyperref[operators]{operator} is unitary if and only if

\begin{gather*}
	\hat{U}^+\hat{U}=\hat{U}\hat{U}^+=I
\end{gather*}

\begin{itemize}
	\tightlist
	\item
	      it preserves inner product
\end{itemize}

\begin{gather*}
	\ket{\alpha '}=\hat{U}\ket{\alpha}\\
	\ket{\beta'}=\hat{U}\ket{\beta}\\
	\braket{\alpha'|\beta'}=\bra{\alpha}\hat{U}^+\hat{U}\ket{\beta}\\
	=\bra{\alpha}I\ket{\beta}\\
	=\braket{\alpha|\beta}
\end{gather*}

\begin{itemize}
	\tightlist
	\item
	      U is basis chagne between orthonormal basis because you can write
	      \(\hat{U}\) as

	      \begin{itemize}
		      \tightlist
		      \item
		            as chatgpt puts it: unitary matrices/operators represent a rotation
		            or change of coordinates in \hyperref[hilbert-spaces]{Hilbert space}, without changing lengths
		            or angles.
	      \end{itemize}
\end{itemize}

\begin{gather*}
	\hat{U}=\hat{U}I
\end{gather*}

I can use \hyperref[completeness]{completeness} relation so that if

\begin{gather*}
	\hat{U}\ket{\alpha_n}=\ket{\beta_n}
\end{gather*}

then we can change basis

\begin{gather*}
	\hat{U}\sum_n\ket{\alpha_n}\bra{\alpha_n}\\
	=\sum_nU\ket{\alpha_n}\bra{\alpha_n}\\
	=\sum_n\ket{\beta_n}\bra{\alpha_n}
\end{gather*}

Both \(\{\ket{\alpha_n}\}\) and \(\{\ket{\beta_n}\}\) are orthonormal
basis
\subsection{Spectral Decomposition}\label{spectral-decomposition}
Let there be a \hyperref[hermitian-operators]{Hermitian operator} \(\hat{A}\) on a finite-dimensional \hyperref[hilbert-spaces]{Hilbert space}.
Spectral theorem says that a  \(\hat{A}\) is unitarily \hyperref[diagonalization]{diagonalizable} and only if it is \hyperref[normal-operator]{normal}.
There exists real eigenvalues \(\{a_n\}\), \hyperref[orthonormality]{orthonormal} \hyperref[eigenvectorseigenvalues]{eigenvectors} \(\{\ket{k}\}\) which form orthogonal \hyperref[projectors]{projectors} \(\{\hat{\Pi}_n\}\) such that

\[
	\hat{A}=\sum_{n=1}^N a_n\hat{\Pi}_n=\sum_{n=1}^N a_n\ket{k}\bra{k}
\]

where projectors satisfy

\begin{enumerate}
	\def\labelenumi{\arabic{enumi}.}
	\tightlist
	\item
	      \(\hat{\Pi}_n\hat{\Pi}_m=0\) for \(n\neq m\) which means they are \hyperref[orthogonality]{orthogonal}
	\item
	      \(\sum_{n=1}^N \hat{\Pi}_n=\hat{I}\) which means they are a complete set of projectors
	\item this projector projects onto the \(a_n\)-\hyperref[eigenspaces]{eigenspace} of \(\hat{A}\)
\end{enumerate}

So \(\{a_n\}\) is the spectrum of \(\hat{A}\).

And that means you can think \(\hat{A}\) as scaling each eigenspace by its own eigenvalue.

\section{Observable}\label{observables}

An observable is represented by a \hyperref[hermitian-matrix]{Hermitian operator} acting on a \hyperref[hilbert-spaces]{Hilbert
	space}.

\section{Pauli Matrices}\label{pauli-matrices}

Three matrices which are defined as

\begin{gather*}
	\sigma^0=\sigma_0=I_2=\begin{pmatrix}1 & 0\\ 0 & 1\end{pmatrix}\\
	\hat{\sigma}_x=\sigma^1=\sigma_1=\begin{pmatrix}
		0 & 1 \\
		1 & 0
	\end{pmatrix}\\
	\hat{\sigma}_y=\sigma^2=\sigma_2=\begin{pmatrix}
		0 & -i \\
		i & 0
	\end{pmatrix}\\
	\hat{\sigma}_z=\sigma^3=\sigma_3=\begin{pmatrix}
		1 & 0  \\
		0 & -1
	\end{pmatrix}
\end{gather*}

Properties

\begin{itemize}
	\tightlist
	\item
	      spacing
\end{itemize}

\begin{gather*}
	\hat{\sigma}^2_i=I_2\quad \forall \quad i\in\{x,y,z\}
\end{gather*}

\begin{itemize}
	\tightlist
	\item
	      commutation
\end{itemize}

note that

The symbol \(\varepsilon_{ijk}\) is the Levi-Civita symbol, an
antisymmetric tensor that encodes signs based on the order of its
indices.

\begin{itemize}
	\tightlist
	\item
	      in 3D
\end{itemize}

\begin{gather*}
	\varepsilon_{ijk}=\begin{cases}
		+1\quad \text{if $(i,j,k)$ is an even permutation of (1,2,3)} \\
		-1\quad \text{if $(i,j,k)$ is an odd permutation of (1,2,3)}  \\
		0\quad\text{if any index repeats}
	\end{cases}
\end{gather*}

\begin{itemize}
	\tightlist
	\item
	      in 2D (only two indices, written \(\varepsilon_{ij}\))
\end{itemize}

\begin{gather*}
	\varepsilon_{ij}=\begin{pmatrix}0 & 1 \\ -1 & 0\end{pmatrix}
\end{gather*}

\begin{gather*}
	[\sigma_i,\sigma_j]=2i\varepsilon_{ijk}\sigma_k
\end{gather*}

where an ``even'' permutation is (1,2,3), (2,3,1), (3,1,2) and an
``odd'' is (2,1,3), (1,3,2), (3,2,1)

\begin{itemize}
	\tightlist
	\item
	      anti commutation
\end{itemize}

Note that the anticommutator is

\begin{gather*}
	\{A,B\}=AB+BA
\end{gather*}

\begin{gather*}
	\{\sigma_i,\sigma_j\}=2\delta_{ij}I_2
\end{gather*}

\begin{itemize}
	\tightlist
	\item
	      completeness
\end{itemize}

\begin{gather*}
	\sigma_0=I_2=\begin{pmatrix}1 & 0\\ 0 & 1\end{pmatrix}
\end{gather*}

the set \(\{\sigma_0,\sigma_x,\sigma_y,\sigma_z\}\) spans all 2x2
complex matrices

\begin{gather*}
	\{I,X,Y,Z\}\quad\text{form a basiss for operator }L(\mathbb{C})^2
\end{gather*}

where

forms a basis for all 2x2 complex matrices. Any 2x2 matrix

\begin{gather*}
	M=\sum_{\mu=0}^3c_\mu\sigma_\mu
\end{gather*}

\begin{itemize}
	\tightlist
	\item
	      whatever this is
\end{itemize}

\begin{gather*}
	\vec{\sigma}=(X,Y,Z)\quad\text{are Hermitian}\\
	\sigma^1=\sigma^x=X\\
	\sigma^2=\sigma^y=Y\\
	\sigma^3=\sigma^z=Z\\
	X^2+Y^2+Z^2=I\\
	tr(X)=tr(Y)=tr(Z)=0\\
	XY=iZ, YZ=iX, ZX=iY\\
	[X, Y] = i\hbar Z, \quad [Y, Z] = i\hbar X, \quad [Z, X] = i\hbar Y\\
	YX=(XY)^+=(iZ)^+=-iZ
\end{gather*}

and

\begin{gather*}
	\hat{A}=a_0\sigma_0+a_1\sigma_x+a_2\sigma_y+a_3\sigma_z\quad a_i\in \mathbb{C}
\end{gather*}

\(\hat{A}\cdot\hat{B}\) can be expanded as

\begin{gather*}
	\hat{A}\hat{B}=\sum_{i=0}^3c_i\sigma^i
\end{gather*}

If \(c_i\in\mathbb{R}\), then \(\hat{A}\) is \hyperref[hermitian-operators]{Hermitian}

\section{Bloch Sphere}\label{bloch-sphere}
A general pure state of a qubit is
\[
	\ket{\psi}=\cos\frac{\theta}{2}\ket{0}+e^{i\phi}\sin\frac{\theta}{2}\ket{1}\quad \theta\in[0,\pi]\quad \phi\in[0,2\pi]
\]
Where the orthogonal states are antipodal (North pole is \(\ket{0}\) and South pole is \(\ket{1}\)).
The equator contains equal superpositions like \(\ket{+}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})\) and \(\ket{-}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})\).

A state can be on the inside of a Bloch sphere which means its a mixed state. If a state is on the surface of the Bloch sphere, it is a pure state.

A rotation by angle \(\theta\) around the axis \(\hat{n}\) is given by
\[
	R_{\hat{n}}(\theta) = e^{-i\theta\,\hat{n}\cdot\vec{\sigma}/2} = \cos\left(\tfrac{\theta}{2}\right)I - i\sin\left(\tfrac{\theta}{2}\right)(n_x\sigma_x + n_y\sigma_y + n_z\sigma_z)
\]
Where
\[
	R_x(\theta) = \begin{pmatrix} \cos\tfrac{\theta}{2} & -i\sin\tfrac{\theta}{2} \\ -i\sin\tfrac{\theta}{2} & \cos\tfrac{\theta}{2} \end{pmatrix}
\]
\[
	R_y(\theta) = \begin{pmatrix} \cos\tfrac{\theta}{2} & -\sin\tfrac{\theta}{2} \\ \sin\tfrac{\theta}{2} & \cos\tfrac{\theta}{2} \end{pmatrix}
\]
\[
	R_z(\theta) = \begin{pmatrix} e^{-i\theta/2} & 0 \\ 0 & e^{i\theta/2} \end{pmatrix}
\]
\section{Observable on a qubit}\label{observable-on-a-qubit}

Any \hyperref[observables]{observable} on a qubit can be written as

\begin{gather*}
	\hat{A}=a_0\sigma_0+\vec{a}\vec{\sigma_{1,2,3}}
\end{gather*}

You can think of \(a_0\sigma_0\) as the scalar value and
\(\vec{a}\vec{\sigma_{1,2,3}}\) as it helps with the \hyperref[bloch-sphere]{Bloch Sphere}.

This can be written in spherical units too

\begin{gather*}
	\hat{n}=(n_x,n_y,n_z)=(\sin\theta\cos\psi,\sin\theta\sin\psi,\cos\theta)
\end{gather*}

spin along \(\hat{n}\) can be described as  \(\hat{S}=\frac{\hbar}{2}\vec{\sigma}\).
This is known as the "spin along the unit vector \(\hat{n}\)".
To do this we project \(\hat{S}\) onto \(\hat{n}\).

\begin{gather*}
	\hat{S}_n=\hat{S}\cdot\hat{n}=\frac{\hbar}{2}\vec{\sigma}\cdot\hat{n}\\
	=\frac{\hbar}{2}\left(n_x\sigma_x+n_y\sigma_y+n_z\sigma_z\right)\\
	=n_x\!\begin{pmatrix}0&1\\1&0\end{pmatrix} + n_y\!\begin{pmatrix}0&-i\\i&0\end{pmatrix} + n_z\!\begin{pmatrix}1&0\\0&-1\end{pmatrix}
\end{gather*}
\[
	= \begin{pmatrix} n_z & n_x - i n_y \\ n_x + i n_y & -n_z \end{pmatrix}
\]
\[
	\Rarr \hat{S}_{\hat{n}} = \frac{\hbar}{2}\begin{pmatrix} n_z & n_x - i n_y \\ n_x + i n_y & -n_z \end{pmatrix}.
\]

We work out the eigenvalues by doing
\[
	det(\hat{S}_{\hat{n}} - \lambda I) = 0
\]
\[
	\Rarr \left(\tfrac{\hbar}{2}n_z-\lambda\right)\left(-\tfrac{\hbar}{2}n_z-\lambda\right) - \left(\tfrac{\hbar}{2}\right)^2(n_x - i n_y)(n_x + i n_y) = 0
\]
\[
	\Rarr \lambda^2 - \left(\tfrac{\hbar}{2}\right)^2(n_x^2+n_y^2+n_z^2) = 0
\]
\[
	\lambda^2=\left(\tfrac{\hbar}{2}\right)^2|\hat{n}|^2\Rarr\lambda=\pm\tfrac{\hbar}{2}|\hat{n}|
\]
Because \(\hat{n}\) is a unit vector,
\[
	\lambda=\pm\tfrac{\hbar}{2}
\]


\section{Degeneracy}\label{degeneracy}
Two or more \hyperref[linear-independence]{linear independent} quantum states (\hyperref[eigenstates]{eigenstates}) share the same \hyperref[eigenvectorseigenvalues]{eigenvalue}.
An energy level \(E_n\) is degenerate if there exist multiple \hyperref[linear-independence]{linearly independent} \hyperref[eigenstates]{eigenstates} \(\ket{\psi_1}, \ket{\psi_2}, \ldots, \ket{\psi_n}\) which satisfy
\[
	\hat{H}\ket{\psi_i}=E_n\ket{\psi_i}\quad \forall i=1,2,\ldots,g
\]
Where if \(g=1\), the energy level is non-degenerate.
If \(g>1\), the energy level is degenerate.
If something is degenerate, that implies there is no unique \hyperref[basis-states]{basis}. What you want to write down should be a \hyperref[projections]{projection} onto the \hyperref[subspaces]{subspaces}.

\[
	\hat{H}\triangleq\begin{pmatrix}
		E_0 &     & 0   \\
		    & E_1 &     \\
		0   &     & E_2
	\end{pmatrix}
	\quad \ket{0}\triangleq\begin{pmatrix}1 \\ 0 \\ 0\end{pmatrix}
	\quad \ket{1a}\triangleq\begin{pmatrix}0 \\ 1 \\ 0\end{pmatrix}
	\quad \ket{1b}\triangleq\begin{pmatrix}0 \\ 0 \\ 1\end{pmatrix}
\]
\[
	\Rarr \hat{H}=E_0\ket{0}\bra{0}+E_1\ket{1a}\bra{1a}+E_2\ket{1b}\bra{1b}
\]
The states \(\ket{1a},\ket{1b}\) are degenerate if and only if \(E_1=E_2\). If \(E_0=E_1, E_1\neq E_2\) then \(\ket{0}\) is degenerate with \(\ket{1a}\).

This just means that you can't detect the difference between states given only the energy/\hyperref[eigenvectorseigenvalues]{eigenvalue}.

Remember the \hyperref[eigenvectorseigenvalues]{eigenvalue} is observed as energy if the  \hyperref[operators]{operator} is the \hyperref[schruxf6dingers-equation-hamiltonian]{Hamiltonian}.

This means the Hamiltonian alone is not enough to determine the state.

This is because if E is degenerate then

\[
	H\ket{\psi_1}= E\ket{\psi_1}
\]
\[
	H\ket{\psi_2}= E\ket{\psi_2}
\]

Now just finding \(E\) doesn't mean we can distinguish the states.

We need another operator \(\hat{A}\) which commutes with \(\hat{H}\) to distinguish the states. Note that just because \(\hat{A}\) commutes with \(\hat{H}\) doesn't mean we can distinguish the states.
\[
	A\ket{\psi_1}= a_1\ket{\psi_1}
\]
\[
	A\ket{\psi_2}= a_2\ket{\psi_2}
\]

And if \(a_1\neq a_2\) then we can distinguish the states.

BUT if \([\hat{A},\hat{H}]\neq0\) then this doesn't always hold true for the entire \hyperref[hilbert-spaces]{Hilbert Space}.

This is because if
\[
	\hat{H}\ket{\psi}= E\ket{\psi}
\]
\[
	\Rarr \hat{A}\hat{H}\ket{\psi}= \hat{A}E\ket{\psi}
\]
\[
	\Rarr (\hat{H}\hat{A}-[\hat{H},\hat{A}])\ket{\psi} = E\hat{A}\ket{\psi}
\]
\[
	\Rarr \hat{H}\hat{A}\ket{\psi} - [\hat{H},\hat{A}]\ket{\psi}= E\hat{A}\ket{\psi}
\]
If \([\hat{H},\hat{A}]=0\) (i.e., H and A commute) then

\[
	\Rarr \hat{H}\hat{A}\ket{\psi} = E\hat{A}\ket{\psi}
\]
This means that \(\hat{A}\ket{\psi}\) is in the \(E\)-\hyperref[eigenspaces]{eigenspace} of \(\hat{H}\). Another way to say this is that \(\hat{A}\ket{\psi}\) is in the set of \hyperref[eigenvectorseigenvalues]{eigenvectors} for H that share \hyperref[eigenvectorseigenvalues]{eigenvalue} E.

This means that \(\hat{A}\) maps a vector in the \(E\)-eigenspace of \(\hat{H}\) to another vector in the \(E\)-eigenspace of \(\hat{H}\).



\section{Compatibility}\label{compatibility}
Compatibility is a property of two \hyperref[operators]{operators}.

If \(\ket{\phi_n}\) is the \(n\)-th \hyperref[eigenstates]{eigenstate} of \(\hat{A}\) (i.e., the \hyperref[eigenbasis]{eigenbasis} of \(\hat{A}\)),

Let
\[
	\hat{A}=\sum_{n=1}^N a_n\ket{\phi_n}\bra{\phi_n}
\]
If \(\hat{B}\) is compatible with \(\hat{A}\) (i.e., \(\hat{B}\) is diagonalizable with \(\hat{A}\)) then,
\[
	\hat{B}=\sum_{n=1}^N b_n\ket{\phi_n}\bra{\phi_n}
\]

\subsection{Compatibility of Observables}\label{compatibility-of-observables}
If two \hyperref[observables]{observables} are compatible, then they can be measured simultaneously with one basic measurement.
This means that observing one doesn't affect the other.

Two \hyperref[operators]{operators} \(\hat{A}\) and \(\hat{B}\) are compatible if and only if they commute, i.e. \([\hat{A},\hat{B}]=0\).

\[
	\hat{A},\hat{B}\quad\text{are compatible}\quad\Leftrightarrow\quad[\hat{A},\hat{B}]=0
\]

Proof
\[
	\Rarr \hat{A}\hat{B}=\sum_{n,m}A_nB_m\ket{\phi_n}\braket{\phi_n|\phi_m}\bra{\phi_m}\\
\]
\[
	=\sum_nA_nB_n\ket{\phi_n}\bra{\phi_n}=\hat{B}\hat{A}\\
\]
\[
	\Larr \hat{A}\ket{\phi_n}=A_n\ket{\phi_n}\\
\]
\[
	\hat{A}\hat{B}\ket{\phi_n}=\hat{B}\hat{A}\ket{\phi_n}=A_n\hat{B}\ket{\phi_n}
\]


Example




\section{Probability Theory}\label{probability-theory}

probability space \(\Sigma\) has points and \hyperref[element]{elements} and associated
probabilities with them

example:

coin flip

\begin{gather*}
	\sum=\{e:H,T\}\rarr\frac{1}{2}
\end{gather*}

Sum of all probability of your space should be 1

\begin{gather*}
	\sum P(\Sigma)=1
\end{gather*}

We have a random variable \(X\). You have an event in that random
variable space called \(x\). The probability of finding that event in
that random variable distribution

\begin{gather*}
	P(X)=\sum_{e\in V_x}P(e)
\end{gather*}

where

\begin{gather*}
	V_x=\{e:X(e)=x\}
\end{gather*}

Note that the probabilities are always between 0 and 1

\begin{gather*}
	0\leq P(x)\leq 1
\end{gather*}

\subsection{Joint Random Variable}\label{joint-random-variables}

We have a joint random variable \((X,Y)\) and we want to find a pair of
values

\begin{gather*}
	P(x,y)=\sum_{x,y\in V_{x,y}}P(e)\\
	V_{x,y}=\{e:X(e)=x, Y(e)=y\}
\end{gather*}

If i have one variable, then the other one can be inferred from the
other.

\begin{gather*}
	P(X)=\sum_yP(x,y)\\
	P(Y)=\sum_xP(x,y)
\end{gather*}

\subsection{Conditional Probability}\label{conditional-probability}

from a joint distribution → if I select \(y\) for \(Y\) with
\(P(Y)\neq 0\)

What is the probability \(x\) given \(y\)

\begin{gather*}
	P(X|Y)=\frac{P(X,Y)}{P(Y)}
\end{gather*}

Special case

If X, Y are independent (if X occurs, Y has no effect, and vice versa)

\begin{gather*}
	P(X,Y)=P(X)P(Y)
\end{gather*}

\subsection{Numerical Random Variable/Expected
	Value}\label{numerical-random-variablesexpected-value}

Uncertainty of a certain measurement

\begin{itemize}
	\tightlist
	\item
	      a numerical RV X labels events by values that are real numbers X.
	\item
	      the mean or expected value
\end{itemize}

\begin{gather*}
	\mathbb{E}[X]=\braket{X}=\sum_X X\cdot P(x)
\end{gather*}

\subsection{Spread}\label{spread}

How do we measure spread from \(\mathbb{E}[X]\)?

Let \(\mu=\mathbb{E}[X]\)

\begin{gather*}
	Var(X)=\braket{(x-\mu)^2}=\braket{X^2}-2\mu\braket{X}+\mu^2\\
	=\braket{X^2}-\braket{X}^2
\end{gather*}

\subsection{Standard Deviation}\label{standard-deviation}

it's the square root of the variance

\begin{gather*}
	\Delta X=\sqrt{Var(X)}=\sqrt{\mathbb{E}[X^2]-\mathbb{E}[X]^2}
\end{gather*}

\subsection{Expected Value of an
	Observable}\label{expected-value-of-observables}

To determine the mean (EV) of an \hyperref[operators]{operator} \(\hat{A}\), we must know the state \(\ket{\psi}\) you're in.
\begin{gather*}
	\mathbb{E}(A)=\sum_nA_nP(n)\\
	=\sum A_n|\braket{n|\psi}|^2\\
	=\sum A_n\braket{\psi|n}\braket{n|\psi}\\
	=\bra{\psi}\sum_n A_n\ket{n}\bra{n}\ket{\psi}\\
	=\bra{\psi}\hat{A}\ket{\psi}\\
	\triangleq\braket{\hat{A}}
\end{gather*}

Another convientn reprsentation is

\begin{gather*}
	\hat{\rho}\triangleq\ket{\psi}\bra{\psi}\\
	\Rarr \braket{\hat{A}}=tr(\hat{\rho}\hat{A})\\
	=tr(\ket{\psi}\bra{\psi}\hat{A})\\
	tr(\bra{\psi}\hat{A}\ket{\psi})
\end{gather*}

where \(tr\) is \hyperref[traces]{trace}
\subsection{Cauchy-Schwarz Inequality}\label{cauchy-schwarz-inequality}
Given two \hyperref[vectors]{vectors} \(\ket{u},\ket{v}\in \mathbb{C}^n\),
\[
	|\braket{u|v}|\leq \|u\|\|v\|
\]
It's squared form can be written as
\[
	|\braket{u|v}|^2\leq \braket{u|u}\braket{v|v}
\]

\section{Born's Rule}\label{borns-rule}

If a \hyperref[quantum-system]{quantum system} is in state \(\ket{\psi}\) and you measure
\hyperref[observables]{observable} \(\hat{A}\) --- the \hyperref[probability-theory]{probability} of getting \hyperref[eigenvectorseigenvalues]{eigenvalue}
\(\lambda_\alpha\) (which corresponds to \hyperref[eigenstates]{eigenstate} \(\ket{\alpha}\)) is

\begin{gather*}
	P_\alpha=|\braket{\alpha|\psi}|^2
\end{gather*}

This only holds true if eigenstate \(\ket{\phi_n}\) is non-degenerate and normalized.
Usually, the convention is that we pick the representative state \(\braket{\phi_n|\phi_n}=1\) (i.e., it is normalized).


Born's rule is a postulate, meaning that it is based on experimental observations and cannot be derived from first principles.

\section{Measuring a Quantum State}\label{measuring-a-quantum-state}
Before:
The system is in a state \(\ket{\psi}\) and you measure the \hyperref[observables]{observable} \(\hat{A}\).

Measure: We get an outcome \(A_n\) with probability \(P_n=\bra{\psi}\hat{\Pi}_n\ket{\psi}\)

Given \hyperref[borns-rule]{Born's rule} for a non-degenerate \hyperref[eigenstates]{eigenstate} \(\ket{\phi_n}\),

\[
	P_n=|\braket{\phi_n|\psi}|^2
\]


This is because \hyperref[borns-rule]{Born's rule}.
Note that \(\hat{\Pi}_n\), the \hyperref[projectors]{projector} onto the \(\lambda_n\)-\hyperref[eigenspaces]{eigenspace} of \(\hat{A}\), is
\[
	\hat{\Pi}_n=\ket{\phi_n}\bra{\phi_n}
\]
\[
	\Rarr \hat{\Pi}_n\ket{\psi}=\ket{\phi_n}\braket{\phi_n|\psi}
\]
\[
	\Rarr =\braket{\phi_n|\psi}\ket{\phi_n}
\]
So
\[
	\Rarr ||\hat{\Pi}_n\ket{\psi}||^2
\]
\[
	\Rarr = ||\braket{\phi_n|\psi}\ket{\phi_n}||^2
\]
\[
	\Rarr = |\braket{\phi_n|\psi}|^2||\ket{\phi_n}||^2
\]
We know that the \hyperref[magnitude]{magnitude} of the eigenstate \(\ket{\phi_n}\) is 1.

After: State collapses to
\[
	\ket{\psi'}=\frac{1}{\sqrt{P_n}}\hat{\Pi}_n\ket{\psi}
\].

The collapse is true only if this is an idealized measurement that is non-destructive.
This means that the measurement must not destroy the system (e.g., photon absorbed by detector). There's no noise as well.

\section{Density Matrix}\label{density-matrix}

Let \(\rho\) be a matrix.

\(\rho\) is a density matrix if it is:

\begin{itemize}
	\tightlist
	\item \hyperref[hermitian-matrix]{Hermitian}
	\item \hyperref[positive-semi-definite-psd-operators]{Positive semi-definite}
	\item \hyperref[traces]{Trace} \(Tr(\rho)=1\)
\end{itemize}

It contains all information needed to predict measurement outcomes.
It is fundementally the probability distribution over quantum states.
For the \hyperref[observables]{observable} \(A\) we see that the \hyperref[expected-value-of-observables]{expected value} is
\[
	\mathbb{E}[A]=tr(\rho A)
\]


\section{Heisenberg Uncertainty Relation}\label{uncertainty-relation}
\[
	\Delta A\Delta B\geq \frac{1}{2}\left|\mathbb{E}\left[[\hat{A},\hat{B}]\right]\right|
\]
\[
	\Delta A^2 \Delta B^2 \geq \left|\bra{\psi}\frac{1}{2j}[\hat{A},\hat{B}]\ket{\psi}\right|^2 = \left(\frac{1}{2j}\mathbb{E}\left[[\hat{A},\hat{B}]\right]\right)^2
\]
Proof
Consider any scalar \(\lambda\in \mathbb{C}\) and any \hyperref[vectors]{vectors} \(\ket{u},\ket{v}\in \mathbb{C}^n\) and we have two states \(\ket{f}\) and \(\ket{g}\).

We subtract it like so \(\ket{f}-\lambda\ket{g}\) is also a vector

Note that for QM systems, \(\braket{u|u}=1\). This is because of \hyperref[borns-rule]{Born's rule} where the number it represents is the total probability, which needs to sum to 1.
\[
	\braket{f-\lambda g|f-\lambda g}\geq 0
\]
\[
	\Rarr\braket{f|f}-\lambda \braket{f|g} - \lambda^*\braket{g|f}+(|\lambda|)^2\braket{g|g}\geq 0
\]

We want to choose the \(\lambda\) that minimizes the expression.

\[
	\lambda=\frac{\braket{g|f}}{\braket{g|g}}\quad \Rarr \quad \lambda^*=\frac{\braket{f|g}}{\braket{g|g}},\quad |\lambda|^2=\frac{\braket{f|g}\braket{g|f}}{\braket{g|g}^2}
\]

Which makes \(\ket{f}-\lambda\ket{g}\) orthogonal to \(\ket{g}\).
\[
	\Rarr\braket{f|f}-\frac{\braket{g|f}}{\braket{g|g}} \braket{f|g} -\frac{\braket{f|g}}{\braket{g|g}}\braket{g|f} +\frac{\braket{f|g}\braket{g|f}}{\braket{g|g}^2}\braket{g|g}\geq 0
\]

Note that \(\braket{g|f}\braket{f|g}=|\braket{f|g}|^2\)

\[
	\Rarr \braket{f|f}-\frac{|\braket{f|g}|^2}{\braket{g|g}}-\frac{|\braket{f|g}|^2}{\braket{g|g}}+\frac{|\braket{f|g}|^2}{\braket{g|g}}\geq 0
\]
\[
	\Rarr \braket{f|f}-\frac{|\braket{f|g}|^2}{\braket{g|g}}\geq 0
\]
Assuming \(\braket{g|g}\geq 0\)
\[
	\Rarr \braket{f|f}\braket{g|g}\geq |\braket{f|g}|^2
\]
This is \hyperref[cauchy-schwarz-inequality]{Cauchy-Schwarz} squared


Consider the \hyperref[observables]{observables} that we defined above. This can be written as
\[
	\ket{f}=(\hat{A}-\mathbb{E}[A]I)\ket{\psi}\quad
	\ket{g}=(\hat{B}-\mathbb{E}[B]I)\ket{\psi}
\]
Note that \(\Delta A\) is the \hyperref[standard-deviation]{standard deviation} of \(\hat{A}\).
\[
	\braket{f|f}=\bra{\psi}(\hat{A}-\mathbb{E}[A]I)\ket{\psi}
\]
\[
	= \bra{\psi}\left(\hat{A}^2-2\mathbb{E}[A]\hat{A}+\mathbb{E}[A]^2I\right)\ket{\psi}
\]
\[
	= \bra{\psi}\hat{A}^2\ket{\psi}-2\mathbb{E}[A]\bra{\psi}\hat{A}\ket{\psi}+\mathbb{E}[A]^2\bra{\psi}I\ket{\psi}
\]
\[
	=\mathbb{E}[\hat{A}^2]-2\mathbb{E}[A]^2+\mathbb{E}[A]^2
\]
\[
	=\mathbb{E}[\hat{A}^2]-\mathbb{E}[A]^2
\]
\[
	=\Delta A^2
\]
By symmetry
\[
	\braket{g|g}=\Delta B^2
\]
Plug into our equation above
\[
	\Delta A^2\Delta B^2\geq |\braket{f|g}|^2
\]
\[
	\Rarr\Delta A^2\Delta B^2\geq (Im\braket{f|g})^2
\]
\[
	\Rarr\Delta A^2\Delta B^2\geq (Im(\bra{\psi}(\hat{A}-\mathbb{E}[A]I)(\hat{B}-\mathbb{E}[B]I)\ket{\psi}))^2
\]
\[
	\Rarr\Delta A^2\Delta B^2\geq (Im(\mathbb{E}[AB]-2\mathbb{E}[A]\mathbb{E}[B]-+ \mathbb{E}[A]\mathbb{E}[B]))^2
\]
\[
	\Rarr\Delta A^2\Delta B^2\geq (Im(\mathbb{E}[AB]-\mathbb{E}[A]\mathbb{E}[B]))^2
\]
Note that \(Im z=\frac{1}{2i}(z-z^*)\)
\[
	\Rarr\Delta A^2\Delta B^2\geq \left(\frac{1}{2i}(\mathbb{E}[AB]-\mathbb{E}[BA])\right)^2
\]
\[
	\Rarr\Delta A^2\Delta B^2\geq \left(\bra{\psi}\frac{1}{2i}[\hat{A},\hat{B}]\ket{\psi}\right)^2
\]

\section{Linear Algebra III}\label{linear-algebra-iii}
Hold on, this is advanced linear algebra!
\subsection{Tensor}\label{tensor}
A scalar is a rank-0 tensor (0D).

A vector is a rank-1 tensor (1D).

A matrix is a rank-2 tensor (2D).

A rank-$n$ tensor is an $n$-dimensional array with $n$ indices.
\subsection{Tensor Product}\label{tensor-product}

Given \hyperref[vector-spaces]{vector spaces} $V$ and $W$, the tensor product $V\otimes W$ is a new vector space satisfying two features:
\begin{enumerate}
	\tightlist
	\item Bilinear pairing where

	      \[
		      \otimes : V\times W\rarr V\otimes W
	      \]
	      satisfies
	      \[
		      (v_1+v_2)\otimes w=v_1\otimes w+v_2\otimes w
	      \]
	      \[
		      v\otimes (w_1+w_2)=v\otimes w_1+v\otimes w_2
	      \]
	      \[
		      (\lambda v)\otimes w=\lambda (v\otimes w)\quad \forall \lambda \in \mathbb{F}
	      \]
	\item For any vector space U and bilinear map $B:V\times W\rarr U$, there exists a unique bilinear map $\tilde{B}:V\otimes W \rarr U$ such that
	      \[
		      B(v,w)=\tilde{B}(v\otimes w)
	      \]

\end{enumerate}

The dimensions are multiplied
\[
	dim(V\otimes W)=dim(V)\times dim(W)
\]

\subsection{Baker Campbell Hausdorff Formula}\label{baker-campbell-hausdorff-formula}

Intelligent set of individuals who worked out that
\[
	e^A e^B = exp(A+B+\frac{1}{2}[A,B]+\frac{1}{12}[A,[A,B]]-\frac{1}{12}[B,[A,B]]+\cdots)
\]

if $[A,B]=0$ does not \hyperref[commutators]{commute}, then $e^A e^B = e^{A+B}$
\subsection{Cross Product}\label{cross-product}
Works on 3-element rank-1 tensors i.e., 3D vectors.
\[
	\vec{A} \times \vec{B} = \begin{vmatrix} \hat{i} & \hat{j} & \hat{k} \\ A_x & A_y & A_z \\ B_x & B_y & B_z \end{vmatrix} = (A_y B_z - A_z B_y)\hat{i} + (A_z B_x - A_x B_z)\hat{j} + (A_x B_y - A_y B_x)\hat{k}
\]
\section{No-Cloning Theorem}\label{no-cloning-theorem}

There exists no unitary operator $U$ that can clone a qubit.
Note that $\otimes$ is the \hyperref[tensor-product]{tensor product}.
\[
	C(\ket{\psi})=\ket{\psi}\otimes\ket{\psi}
\]

Proof

Let
\[
	C\ket{0}\rarr \ket{0}\ket{0}
\]
\[
	C\ket{1}\rarr \ket{1}\ket{1}
\]

Then

\[
	C\ket{+}\rarr=C\frac{\ket{0}+\ket{1}}{2}\\
\]
\[
	\neq\frac{1}{\sqrt{2}}(\ket{0}\ket{0}+\ket{1}\ket{1})\\
\]
\[
	\neq \ket{+}\ket{+}
\]
This breaks the \hyperref[operators]{linearity of operators}.

\section{Quantum Information Capacity}\label{quantum-information-capacity}

A classical bit has 2 states: 0 and 1.
\[
	\{0,1\}
\]

A qubit has infinite distinct states. Linear combination (superposition) of
\[
	\ket{\psi} = \alpha\ket{0} + \beta\ket{1}
\]

Example

We can let there be three states in \hyperref[quantum-system]{spin-1/2}.
\[
	\ket{z+}\quad\text{british coming by land}
\]
\[
	\ket{z-}\quad\text{british coming by sea}
\]
\[
	\ket{x+}\quad\text{british not coming}
\]

So i can measure \(S_z\) to be either \(+\hbar/2\) or \(-\hbar/2\).
If it is \(+\hbar/2\), then the british are not coming by sea. If it is \(-\hbar/2\), then the british are not coming by land.

I can also measure \(S_x\) to be either \(+\hbar/2\) or \(-\hbar/2\).
If it is \(+\hbar/2\), we don't get any extra information as it could equally be either three. If it is \(-\hbar/2\), then the british are coming either by land or sea.

How reliably can we distinguish between a set of quantum states that serve as signals?

Consider Alice and Bob. Alice sends quantum state to Bob and this state is living in some \hyperref[hilbert-spaces]{Hilbert space} \(H\)
\[
	\ket{\psi}\in H\quad dim(H)=d
\]

Alice encodes \(N\) equally likely messages
\[
	\{\ket{\alpha}\}_{\alpha=1}^N
\]

Bob makes measurements in some \hyperref[basis-states]{basis}
\[
	\{\ket{k}\}_{k=1}^{d'}
\]

Bob measures a larger Hilbert space \(H'\) because he doesn't know the Hilbert space Alice is using.
\[
	H\subset H'\quad dim(H')=d'>d
\]
Bob still needs to make a determination. But he still needs to make a decision on which of the \(N\) messages sent.

Let
\[
	k=1,2,3,4,\cdots,d'\quad d'>N
\]

for some of these outcomes you can get maybe you assign some like $1,2$ to be message 1 $C_1$ and $3$ to message 2 $C_2$, etc. then you need to cover all possibilities of the message Alice sends. You want to map the outcome to messages. Some of the bins may be empty.

The success probability for Bob to correctly distinguish the different messages.
\[
	P_s=\sum_{\alpha=1}^NP(\alpha)\sum_{k\in C_\alpha}P(outcome~k|\text{message $\alpha$ was sent})\\
\]
Recall \hyperref[borns-rule]{Born's rule} that
\[
	=\sum_{\alpha =1}^N\frac{1}{N}\sum_{k\in C_\alpha}|\braket{k|\alpha}|^2
\]
We want to try to upperbound this. Let $\{\ket{\phi_n}\}_{r=1}^d$ be some \hyperref[orthonormality]{orthonormal basis} of $H$

We know that
\[
	\Pi\ket{\alpha}=\ket{\alpha}
\]
And also
\[
	(\Pi\ket{\alpha})^\dag=\ket{\alpha}^\dagger
\]
\[
	\bra{\alpha}\Pi^\dag=\bra{\alpha}
\]
\[
	\bra{\alpha}\Pi=\bra{\alpha}
\]
So
\[
	P_s=\frac{1}{N}\sum_\alpha\sum_{k\in C_\alpha}\braket{\alpha|k}\braket{k|\alpha}\\
\]
\[
	=\frac{1}{N}\sum_\alpha\sum_{k\in C_\alpha}\braket{\alpha|\Pi|k}\braket{k|\Pi|\alpha}\\
\]
\[
	=\frac{1}{N}\sum_\alpha\sum_{k\in C_\alpha}\bra{\alpha}(\Pi\ket{k}\bra{k}\Pi)\ket{\alpha}
\]
\[
	= \frac{1}{N}\sum_\alpha \sum_{k\in C_\alpha}tr(\bra{\alpha}\Pi\ket{k}\bra{k}\Pi\ket{\alpha})\\
\]
Using \hyperref[traces]{trace} cyclic property
\[
	= \frac{1}{N}\sum_\alpha \sum_{k\in C_\alpha}tr(\ket{\alpha}\bra{\alpha}\cdot \Pi\ket{k}\bra{k}\Pi)\\
\]
\[
	\leq \frac{1}{N}\sum_\alpha \sum_{k\in C_\alpha}tr(\Pi\ket{k}\bra{k}\Pi)\\
\]
\[
	\leq \frac{1}{N}\sum_\alpha \sum_{k\in C_\alpha}tr(\Pi^2\cdot\ket{k}\bra{k})\\
\]
\[
	\leq \frac{1}{N}tr(\Pi^2\sum_\alpha\sum_{k\in C_\alpha}\ket{k}\bra{k})\\
\]
\[
	\leq \frac{1}{N}tr(\Pi^2I)\\
\]
\[
	\leq\frac{1}{N}tr(\Pi)\\
\]
Using \hyperref[traces]{trace} cyclic property and the fact that \(tr(\braket{\phi_r|\phi_r})=\braket{\phi_r|\phi_r}\) as it is a \(1\times 1\) matrix.
\[
	\leq\frac{1}{N}tr(\sum_{r=1}^d\ket{\phi_r}\bra{\phi_r})
\]
\[
	\leq\frac{1}{N}\sum_{r=1}^dtr(\braket{\phi_r|\phi_r})
\]
\[
	\leq\frac{1}{N}\sum_{r=1}^d\braket{\phi_r|\phi_r}
\]
\[
	\leq\frac{1}{N}\sum_{r=1}^d1
\]
\[
	\leq\frac{d}{N}
\]
This means the probability of a successful decoding is at most the number of dimensions of the Hilbert space divided by the number of possible messages Alice can send

\section{Basic Decoding Theory}\label{basic-decoding-theory}

This is based on \hyperref[quantum-information-capacity]{quantum information capacity}.
Given that Alice is trying to send a message to Bob, the decoding error is
\[
	P_E=1-P_s\geq 1-\frac{d}{N}
\]
Note that
\begin{enumerate}
	\item If \(N\geq d\), then \(P_E\geq 0\) -- there is no way to reliably distinguish messages
	\item lower bound in theorem also holds even if alice sends mixed states \(\{\rho_\alpha\}_{\alpha=1}^N\)
	\item A quantum system with \hyperref[hilbert-spaces]{Hilbert space} dimension \(d\) has information capacity of \(\log_2(d)\) bits. This is the maximum number of bits that can be stored, communicated, and read reliably. This is the same amount of infromation from classical d-state system.
\end{enumerate}

\section{Basic Distinguishability Theory}\label{basic-distinguishability-theory}

This refines \hyperref[basic-decoding-theory]{basic decoding theory}.

Probability of successfully distinguishing between two equally likely pure states \(\ket{\alpha_0}, \ket{\alpha_1}\) where \(\braket{\alpha_0|\alpha_1}=\cos\theta\) is
\[
	P_s\leq \frac{1}{2}(1+\sin\theta)
\]

Example I

We want to find the optimal measurement to distinguish between two equally likely pure states \(\ket{\alpha_0}, \ket{\alpha_1}\) where \(\braket{\alpha_0|\alpha_1}=\cos\theta\).
We can try to do

\[
	\ket{\alpha_0}=\ket{0}\quad\ket{\alpha_1}=\ket{+}
\]
\[
	\cos\theta =\sin\theta=\frac{1}{\sqrt{2}}
\]
and there are is a $\ket{1}$ basis that is ortho to $\ket{0}$. If we measure and get $\ket{1}$ it couldn’t have come from $\ket{\alpha_0}$

But it isn't optimal

Example II

The optimal way is to get the bisection angle between $\ket{\alpha_1},\ket{\alpha_0}$
\[
	P_s(\tau)=P(\alpha_0)P(\text{measure $k_0$ given $\ket{\alpha_0}$})+P(\alpha_1)P(\text{measure $k_1$ given $\ket{\alpha_1}$})\\
\]
\[
	=\frac{1}{2}\cos^2(\tau)+\frac{1}{2}\cos^2(\frac{\pi}{2}-\tau-\theta)
\]
Optimize to get
\[
	\tau^*=\frac{1}{2}\left(\frac{\pi}{2}-\theta\right)\\
\]
\[
	\Rarr \max P_s(\tau)=\frac{1}{2}\cos^2\left(\frac{\pi}{4}-\frac{\theta}{2}\right)+\frac{1}{2}\cos^2(\frac{\pi}{4}-\frac{\theta}{2})\\
\]
\[
	=\frac{1}{2}\left(1+\sin\theta\right)
\]

\section{Quantum Cryptography}\label{quantum-cryptography}

\subsection{Classical Cryptography}\label{classical-cryptography}
How do you communicate securely over a public channel?

Using a shared secret key! (Symmetric key encryption)

Let ciphertext $C$, plaintext $P$, and key $K$ be \hyperref[strings]{strings} of bits $\{0,1\}^*$.
Let $E$ be an encryption \hyperref[functions]{function} and $D$ be a decryption \hyperref[functions]{function}.

\[
	C=E(P, K)
\]
\[
	P=D(C, K)
\]

There are three people Alice, Bob and Eve where Alice is trying to send data to Bob over a public channel and Eve is listening to it and maliciously trying to obtain the data.

How do we securely share the secret key $K$ in the first place? Classically we use RSA or Diffie-Hellman, but this relies on
assumptions about hardness of a computational problem, which may not be hard for a quantum coputer.


In the quantum world, we use quantum key distribution (QKD) to securely share the secret key $K$ without Eve being able to intercept it.

\subsection{BB84 Protocol}\label{bb84-protocol}
This is a type of quantum key distribution protocol. It is named after Bennett and Brassard in 1984.

Alice chooses a basis $b$ (subset of the \hyperref[pauli-matrices]{Pauli matrices}) and a bit $p$ (i.e., parent)
\[
	b\in\{X,Z\}
	\quad
	p\in\{0,1\}
\]

We know there are an infinite number of possible bases in a qubit. There exists three mutually unbiased bases in a qubit $\{Z,X,Y\}$. BB84 chooses $\{X,Z\}$ as part of the definition of the protocol.

Because a qubit is $d=2$ by definition, we can only choose from 2 parents.

Alice then creates a qubit state $\ket{\psi_{b,p}}$ to send to Bob.
BB84 defines the following mapping:

\begin{tabular}{|c|c|l|}
	\hline
	basis $b$ & parent $p$ & qubit state $|\psi_{b,p}\rangle$                         \\
	\hline
	$Z$       & $0$        & $|0\rangle$                                              \\
	$Z$       & $1$        & $|1\rangle$                                              \\
	$X$       & $0$        & $|+\rangle = \tfrac{1}{\sqrt{2}}(|0\rangle + |1\rangle)$ \\
	$X$       & $1$        & $|-\rangle = \tfrac{1}{\sqrt{2}}(|0\rangle - |1\rangle)$ \\
	\hline
\end{tabular}

Example

To generate a shared key, Alice chooses a string of random bases $\in\{X,Z\}^n$ and parents $\in\{0,1\}^n$ to create a string of qubit $\ket{\psi_{b,p}}$ to send to Bob. Where $n$ is the number of bits to send.

% Step 1: Alice generates two random strings and sends qubits
\begin{tabular}{|c|cccccccc|}
	\hline
	basis $b$                  & $X$         & $Z$         & $Z$         & $X$         & $Z$         & $X$         & $X$         & $Z$         \\
	parent $p$                 & $1$         & $0$         & $1$         & $1$         & $0$         & $1$         & $0$         & $1$         \\
	qubit $|\psi_{b,p}\rangle$ & $|-\rangle$ & $|0\rangle$ & $|1\rangle$ & $|-\rangle$ & $|0\rangle$ & $|-\rangle$ & $|+\rangle$ & $|1\rangle$ \\
	\hline
\end{tabular}

Bob chooses a string of random bases $\in\{X,Z\}^n$ and measures the $i$-th qubit Alice sent in the $i$-th base.

% Step 2: Bob measures in random bases from {X, Z}
\begin{tabular}{|c|cccccccc|}
	\hline
	Bob's basis $b'$    & $X$         & $X$         & $Z$         & $Z$         & $X$         & $Z$         & $X$         & $Z$         \\
	\hline
	example result      & $|-\rangle$ & $|-\rangle$ & $|1\rangle$ & $|0\rangle$ & $|+\rangle$ & $|1\rangle$ & $|+\rangle$ & $|1\rangle$ \\
	\quad or            & $|-\rangle$ & $|+\rangle$ & $|1\rangle$ & $|1\rangle$ & $|-\rangle$ & $|0\rangle$ & $|+\rangle$ & $|1\rangle$ \\
	\hline
	inferred parent (?) & $1$         & $1$         & $1$         & $0$         & $0$         & $1$         & $0$         & $1$         \\
	correct basis       & \checkmark  &             & \checkmark  &             &             &             & \checkmark  & \checkmark  \\
	secret key          & $1$         &             & $1$         &             &             &             & $0$         & $1$         \\
	\hline
\end{tabular}

Alice and Bob both announce their basis but not their measurement results. Correct basis for half the qubits. Then use the subsequence as the secret key $K$.

\[
	K = 1101
\]

Alice and Bob then agree on a random subset $T\subset K$. They publicly announce their bit values only at positions in $T$.

\[
	QBER=\frac{\text{number of pos. in $T$ where bits differ}}{|T|}
\]

If quantum bit error rate QBER is lower than the threshold, then proceed with the remaining bits not in $T$. In BB84, the threshold is typically around 11\% because of Shannon-theoretic entropy.

This is secure against Eve as if Eve intercepts a qubit, Eve can only distinguish among $\{\ket{0},\ket{1},\ket{+},\ket{-}\}$
where error probability  $P_E\geq 1-\frac{d}{N}$ given in \hyperref[basic-decoding-theory]{basic decoding theory}.

If Eve sends qubit along to Bob, they would be wrong half of the time. Alice and Bob can find out in the testing stage. During testing stage, some bits of the key are sent between Alice and Bob and discarded if they do not match. This reduces information that Eve has about the secret key. The probability of the bit disagreeing given Eve intercepted is $\frac{1}{2}\times\frac{1}{2}=\frac{1}{4}$ whereas if Eve doesn't intercept, the probability of the bit disagreeing is $0$.

Due to the \hyperref[no-cloning-theorem]{no-cloning theorem}, Eve cannot copy the qubit and send it to Bob.



\section{Quantum Dynamics}\label{quantum-dynamics}
Quantum dynamics is the study of how quantum systems evolve over time.
\subsection{Isolated System}\label{isolated-system}

This builds on \hyperref[schruxf6dingers-equation-hamiltonian]{Schrödinger's equation}.
State changes from $t_1\rarr t_2$ according to some linear \hyperref[operators]{operator} $U(t_2, t_1)$.
\[
	\ket{\psi(t_2)}=U(t_2, t_1)\ket{\psi(t_1)}
\]

Properties
\begin{enumerate}
	\tightlist
	\item $t_2=t_1$ gives no change
	      \[
		      \ket{\psi(t_1)}=U(t_1, t_1)\ket{\psi(t_1)}
	      \]
	      \[
		      U(t_1, t_1)=I
	      \]
	\item composition
	      \[
		      \ket{\psi(t_3)}=U(t_3, t_2)\ket{\psi(t_2)}
	      \]
	      \[
		      =U(t_3, t_2)U(t_2, t_1)\ket{\psi(t_1)}
	      \]
	      \[
		      =U(t_3, t_1)\ket{\psi(t_1)}
	      \]
	      hence
	      \[
		      U(t_3, t_1)=U(t_3, t_2)U(t_2, t_1)
	      \]
	\item inverse
	      \[
		      I=U(t_1,t_2)U(t_2,t_1)
	      \]
	      \[
		      U(t_1,t_2)=U^{-1}(t_2,t_1)=U^\dag(t_2, t_1)
	      \]
	\item diagonalizable
	      \[
		      U^\dag U = U U^\dag
	      \]
	      this means \hyperref[spectral-decomposition]{spectral theorem} applies.
	      For every orthonormal basis $\{\ket{k}\}$,
	      \[
		      U=\sum_k\lambda_k\ket{k}\bra{k}
	      \]
	      \[
		      I=\bra{k}U^\dag U\ket{k}
	      \]
	      \[
		      \lambda_k=e^{i\theta_k}
	      \]
\end{enumerate}


\subsection{Unitary Evolution}\label{unitary-evolution}

Postulates of an \hyperref[isolated-system]{isolated system}:

\begin{enumerate}
	\tightlist
	\item $U(t_2,t_1)$ should be independent of initial state.
	      \[
		      \alpha\ket{\psi}+\beta\ket{\phi}\rarr U(t_2,t_1)(\alpha\ket{\psi}+\beta\ket{\phi})=
	      \]
	\item must preserve \hyperref[probability-theory]{probablity} (must sum to 1) if isolated which means $U$ must be \hyperref[unitary-operator]{unitary} if \hyperref[hilbert-spaces]{Hilbert space} is finite dimensional.
	\item assume $U$ is unitary in infinite dimensional Hilbert space.
\end{enumerate}

Unpacking 2. we see that
\[
	\braket{\psi|\psi}=1
\]
This is because \hyperref[borns-rule]{Born's rule} says that, over $\ket{k}$ basis states,
\[
	\sum_k |\braket{k|\psi}|^2=1
\]
as probability must sum to 1.
Using \hyperref[absolute-value-of-complex-number]{this lemma}, we get
\[
	\sum_k|\braket{k|\psi}|^2=\sum_k\braket{\psi|k}\braket{k|\psi}=\bra{\psi}\left(\sum_k\ket{k}\bra{k}\right)\ket{\psi}
\]
\[
	=\bra{\psi}I\ket{\psi}=\braket{\psi|\psi}
\]

so if we evolve it over time then
\[
	(U\ket{\psi})^\dag U\ket{\psi}=1\quad\forall\quad\ket{\psi}
\]
\[
	\Rarr \quad \bra{\psi}U^+U \ket{\psi}=1
\]
\[
	\Rarr \quad U^+U=I
\]

This means that $U$ is \hyperref[unitary-operator]{unitary}.

Note that

\[
	U(t_0,t)=U^{-1}(t,t_0)=U^\dag(t,t_0)
\]
\subsection{Schrödinger Equation, Hamiltonian II}\label{schrodinger-equation}
Let's derive the Schrödinger equation from the postulates of an \hyperref[isolated-system]{isolated system}.

Let $t_0$ be the initial time and $t$ be the time after some time has passed.

Let
\[
	|\psi(t)\rangle = U(t, t_0) |\psi(t_0)\rangle
\]
So

\[
	\frac{\partial}{\partial t} |\psi(t)\rangle = \frac{\partial U(t, t_0)}{\partial t} |\psi(t_0)\rangle
\]
Due to its inverse property where $U(t_0,t)U(t,t_0)=I$,

\[
	= \frac{\partial U(t, t_0)}{\partial t} \, U(t_0, t) |\psi(t)\rangle
\]

\[
	= \underbrace{\frac{\partial U(t, t_0)}{\partial t} \, U^\dagger(t, t_0)}_{\Lambda(t, t_0)} |\psi(t)\rangle
\]


Properties
\begin{enumerate}
	\tightlist
	\item \(\Lambda(t, t_0)\) is \hyperref[anti-hermitian-operators]{anti-Hermitian}.

	      Note:
	      \[
		      \Lambda^\dagger(t, t_0) = U(t, t_0) \, \frac{\partial U^\dagger(t, t_0)}{\partial t}
	      \]

	      Differentiate \(U(t, t_0) U^\dagger(t, t_0) = I\):
	      \[
		      \frac{\partial}{\partial t}\!\left[U(t, t_0) U^\dagger(t, t_0)\right] = \frac{\partial U(t, t_0)}{\partial t} U^\dagger(t, t_0) + U(t, t_0) \frac{\partial U^\dagger(t, t_0)}{\partial t} = 0
	      \]

	      \[
		      \Rightarrow \;\; \Lambda(t, t_0) + \Lambda^\dagger(t, t_0) = 0
	      \]

	      \[
		      \Rightarrow \;\; \Lambda = -\Lambda^\dagger \;\;\text{is anti-Hermitian.}
	      \]
	\item \(\Lambda(t, t_0)\) is independent of $t_0$
	      \[
		      \Lambda (t,t_0)=\frac{\partial U(t, t_0)}{\partial t} U^\dagger(t, t_0)
	      \]
	      \[
		      =\frac{\partial U(t, t_0)}{\partial t} U(t_0,t_1)U^\dag(t_0,t_1)
		      U^\dagger(t, t_0)
	      \]
	      \[
		      =\frac{\partial}{\partial t}[U(t,t_0)U(t_0,t_1)][U(t,t_0)U(t_0, t_1)]^\dag
	      \]
	      \[
		      =\frac{\partial U(t,t_1)}{\partial t}U^\dag(t,t_1)=\Lambda (t,t_1)
	      \]
	      For any $t_0,t_1$
\end{enumerate}

Let

\[
	\Lambda (t)\triangleq \Lambda (t,t_0)
\]
\[
	\Rarr \frac{d}{dt}\ket{\psi(t)} = \frac{\partial}{\partial t}U(t, t_0)U^\dag(t,t_0) |\psi(t)\rangle
\]
\[
	=\Lambda (t)\ket{\psi(t)}
\]

Then, let
\[
	\boxed{H(t)\triangleq i\hbar \Lambda (t) = i\hbar \frac{\partial U(t, t_0)}{\partial t} U^\dagger(t, t_0)}
\]

This is known as the Hamiltonian, which has the units of energy. In natural units where $\hbar=1$, it is simply inverse time $s^{-1}$.

We arrive at the Schrödinger equation:
\[
	\boxed{i\hbar \frac{d}{dt}\ket{\psi(t)}=H(t)\ket{\psi(t)}}
\]
Remarks

\begin{enumerate}
	\tightlist
	\item Derived Schrödinger equation from only the assumption of unitary time evolution. The Hamiltonian operator is derived from $U(t,t_0)$
	\item Converse is true where we can construct $U(t,t_0)$ from knowledge of $H(t)$
	\item in many physical systems, it easier to find $H(t)$ than finding $U(t,t_0)$
\end{enumerate}

\subsection{Example of Finding Hamiltonian}\label{example-of-finding-hamiltonian}

Let there be a particle in space where energy is kinetic energy plus potential energy.

In classical physics, we know that
\[
	KE = \frac{1}{2}mv^2 = \frac{p^2}{2m}
\]
where $KE$ is kinetic energy, $m$ is mass, $v$ is velocity and $p=mv$ is momentum.
We know that potential energy $V(x)$ is a function of position $x$.

Hence in classical physics, energy is
\[
	E=KE+V(x)=\frac{p^2}{2m}+V(x)
\]

A wavelength $\lambda$ travelling in the direction $x$ can be written as $e^{ikx}$ where wave number $k$ is defined as
\[
	k\triangleq \frac{2\pi}{\lambda}
\]

de Broglie in 1924 hypothesized that every particle has wave-like behavior with wavelength $\lambda$
\[
	\lambda = \frac{h}{p}
\]
\[
	\Rarr p=\frac{h}{2\pi/k}=\hbar k
\]

But we want this as an operator where the plane wave is an \hyperref[eigenstates]{eigenstate} of \(\hat{p}\) with eigenvalue $\hbar k$ (its momentum).

\[
	\hat{p}e^{ikx}=\hbar ke^{ikx}
\]
Now differentiate $e^{ikx}$ w.r.t $x$
\[
	\Rarr \frac{\partial}{\partial x}e^{ikx}=ike^{ikx}
\]
Multiply by $\hbar/i$
\[
	\Rarr \frac{\hbar}{i}\frac{\partial}{\partial x}e^{ikx}=\hbar ke^{ikx}
\]
Comparing like terms we find
\[
	\hat{p}=\frac{\hbar}{i}\frac{\partial}{\partial x}
\]


Anyways, after doing all of that, we're going to guess the Hamiltonian.
Guess that the Hamiltonian is $H(t)=E$
\[
	H(t)=\frac{\hat{p}^2}{2m}+V(\hat{x})
\]
Expand the state into basis via \hyperref[spectral-decomposition]{spectral decomposition}

\[
	\ket{\psi(t)}=\sum_k\braket{x_k|\psi(t)}\ket{x_k}=\int dx \braket{x|\psi(t)}\ket{x}
\]

The intergral is the continuum limit of the discretized version on the left.
Let's take the Schrödinger equation and apply it to the state.

\[
	\bra{x}i\hbar \frac{d}{dt}\ket{\psi(t)}=\bra{x}H(t)\ket{\psi(t)}
\]
LHS and RHS are both manipulated to get
\[
	i\hbar \frac{\partial}{\partial t}\braket{x|\psi(t)} =H(t)\braket{x|\psi(t)}
\]
\[
	\Rarr i\hbar \frac{\partial}{\partial t}\braket{x|\psi(t)} =\bra{x}\frac{\hat{p}^2}{2m}\ket{\psi(t)}+V(x)\braket{x|\psi(t)}
\]
\[
	\Rarr i\hbar \frac{\partial}{\partial t}\braket{x|\psi(t)} =-\frac{\hbar^2}{2m}\frac{\partial^2}{\partial x^2}\braket{x|\psi(t)}+V(x)\braket{x|\psi(t)}
\]
\[
	\Rarr i\hbar \frac{\partial}{\partial t}\braket{x|\psi(t)} =\left(-\frac{\hbar^2}{2m}\frac{\partial^2}{\partial x^2}+V(x)\right)\braket{x|\psi(t)}
\]
If we solve this for specific systems and check against measurements, it will confirm whether or not our guess for $H(t)$ was right
\subsection{Wavefunction}\label{wavefunction}
Where $\ket{\psi(t)}$ is a \hyperref[ket-states]{state} in the \hyperref[hilbert-spaces]{Hilbert space}, $\psi(x,t)$ is the wavefunction w.r.t time $t$ and position $x$.

Let
\[
	\psi(x,t)\triangleq\braket{x|\psi(t)}
\]
\subsection{Uniform dynamics}\label{uniform-dynamics}

This is a time-indepndent Hamiltonian $H(t)=H$.

We set an ODE for $U$ which we derived earlier.

\[i\hbar\frac{dU}{dt} = HU \iff \frac{dU}{dt} = -\frac{iH}{\hbar}U\]

If we try it for a scalar value, it is similar to
\[
	\frac{dy}{dt} = ky(t) \implies y(t) = e^{kt}y(0)
\]

Recall \hyperref[matrix-exponentials]{matrix exponential}:

Guess that $U(t,t_0) = e^{-iHt/\hbar}U_0$

Check:
\[\frac{dU}{dt} = -\frac{iH}{\hbar}e^{-iHt/\hbar}U_0 = -\frac{iH}{\hbar}U \]

Hence true.

Note $U_0$ is some integration constant which we haven't defined yet.
We find $U_0$

\[I = U(t_0,t_0) = e^{-iHt_0/\hbar}U_0 \implies U_0 = e^{iHt_0/\hbar}\]

\[\implies U(t,t_0) = e^{-iHt/\hbar}e^{iHt_0/\hbar}\]

via \hyperref[baker-campbell-hausdorff-formula]{Baker Campbell Hausdorff Formula}, since $[H,H]=0$ (trivially as any operator commutes with itself)
\[
	U(t,t_0) = e^{-iHt/\hbar}e^{iHt_0/\hbar}
\]
\[
	\Rarr\boxed{U(t,t_0) = e^{-iH(t-t_0)/\hbar}}
\]

Hence this solves SE
\[
	i\hbar \frac{d}{dt}U(t,t_0) = HU(t,t_0)
\]
\subsection{Heisenberg Picture}\label{heisenberg-picture}

Let observable $\hat{A}$. The Heisen picture operator is
\[
	\tilde{A}(t)=U^\dag (t)AU(t)
\]

\begin{tabular}{|l|c|c|}
	\hline
	           & Schrödinger Picture                                     & Heisenberg Picture                                                      \\
	\hline
	State      & $|\psi(t)\rangle = U(t)|\psi(0)\rangle$                 & $|\psi(0)\rangle$ constant                                              \\
	Observable & $A$ constant                                            & $\tilde{A}(t) = U^\dagger(t)\, A\, U(t)$                                \\
	Dynamics   & $i\hbar\dfrac{d|\psi\rangle}{dt} = H(t)|\psi(t)\rangle$ & $\dfrac{d\tilde{A}}{dt} = \dfrac{i}{\hbar}[\tilde{H}(t), \tilde{A}(t)]$ \\
	\hline
\end{tabular}

For \hyperref[uniform-dynamics]{uniform dynamics}, $\tilde{H}(t)=\tilde{H}$. That means $U(t)\triangleq U(t,t_0)= e^{-iHt/\hbar}e^{iHt_0/\hbar}$

We know that something \hyperref[commutators]{commutes} with itself, so $[U(t), H]=0$


\[
	\Rarr \tilde{H}(t)=U^\dag(t)HU(t)=H
\]

Heisenberg equation of motion for $\tilde{A}(t)$

\[
	\frac{d}{dt}A(t)=\frac{d}{dt}(U^\dag AU + U^\dag A\frac{dU}{dt})
\]
\[
	=-\frac{1}{i\hbar}U^\dag H A U+\frac{1}{i\hbar}U^\dag A H U
\]
\[
	=-\frac{i}{\hbar}(U^\dag HU U^\dag A U - U^\dag A U U^\dag H U)
\]
because \hyperref[commutators]{commutator},
\[
	\boxed{\frac{d}{dt}\hat{A}(t)=\frac{i}{\hbar}[\hat{H}(t),\hat{A}(t)]}
\]

\subsection{Ehrenfest Theorem}\label{observables-in-time-evolution}
Given \hyperref[observables]{observable} $\hat{A}$, recall the \hyperref[expected-value-of-observables]{expectation value} of $\hat{A}$ is
\[
	\mathbb{E}[A(t)]=\bra{\psi(t)}A(t)\ket{\psi(t)}
\]
If we \hyperref[differentiation]{differentiate} w.r.t time
\[
	\frac{d}{dt}\mathbb{E}[A(t)]=\left(\frac{d}{dt}\bra{\psi(t)}\right)A(t)\ket{\psi(t)}+\bra{\psi(t)}A(t)\left(\frac{d}{dt}\ket{\psi(t)}\right)
\]
\[
	= -\frac{1}{i\hbar}\bra{\psi}HA\ket{\psi}+\bra{\psi}A\frac{1}{i\hbar}H\ket{\psi}
\]
because \hyperref[commutators]{commutator},
\[
	=\frac{i}{\hbar}\bra{\psi}[H,A]\ket{\psi}
\]


Ehrenfest theorem

\[
	\boxed{\frac{d}{dt}\mathbb{E}[A(t)]=\frac{i}{\hbar}\mathbb{E}\left[[H,A]\right]}
\]

Remarks

\begin{enumerate}
	\tightlist
	\item If $[H,A]=0$, then $\frac{d}{dt}\mathbb{E}[A(t)]=0$
	      The \hyperref[expected-value-of-observables]{expected value} of A is conserved if A and H are \hyperref[compatibility]{compatible}. For example, energy H is always conserved in an \hyperref[isolated-system]{isolated system}!
	\item Sometimes convientn to change basis to Heisenberg picture to study observable dynamics
\end{enumerate}




\section{Example of Quantum Dynamics}\label{example-of-quantum-dynamics}
This depends on \hyperref[quantum-dynamics]{quantum dynamics}.

\subsection{Example 1}\label{example-1-of-quantum-dynamics}
An electron is in an magnetic field in the $z+$ direction. There is a particle in state
\[
	\ket{\psi(0)}=\begin{pmatrix}\cos\theta_0/2 \\ e^{i\phi_0}\sin\theta_0/2\end{pmatrix}
\]
Find out the expected value of $S_z, S_x, S_y$ at time $t$.

Classical charged particle has
\[\vec{\mu}=\frac{q}{2m}\vec{s}\]

A quantum one also has a $g$-factor
\[\vec{\mu}=g\frac{q}{2m}\vec{s}\]

In the case of an electron, $q=e, m=m_e, g\approx 2$
\[
	\vec{\mu}=g\underbrace{\frac{e\hbar}{2m_e}}_{\mu_B}\frac{1}{\hbar}\vec{s}
\]
Where $\mu_B$ is the Bohr magneton.

The spin in magnetic field $\vec{B}$ experiences a torque.

Clasically, the energy of a magnetic dipole in a B field is
\[
	U=-\vec{\mu}\cdot\vec{B}
\]

In quantum mechanics, the energy is the Hamiltonian operator $\hat{H}$
Let
\[
	\gamma = \frac{g\mu_B}{\hbar}
\]
So
\[
	\hat{H}=-\vec{\mu}\cdot\vec{B}=-\gamma\vec{s}\cdot\vec{B}
\]

Since B field is in the $z+$ direction, $\vec{B}=B\hat{S_z}$

\[
	\hat{H}=-\gamma B \frac{\hbar}{2}\begin{pmatrix}1 & 0 \\ 0 & -1\end{pmatrix}=-\gamma B \hat{S}_z
\]

To do it the Schrödinger equation way, take the eigenvalues and \hyperref[eigenstates]{eigenstates} of $H$
\[
	H\ket{z\pm}=\mp\gamma B \frac{\hbar}{2}\ket{z\pm}
\]
Because Hamiltonian is time-independent, we can use the \hyperref[uniform-dynamics]{uniform dynamics} solution to find $U(t,t_0)$
\[
	U(t,0)\ket{z\pm}=e^{-iHt/\hbar}\ket{z\pm}=e^{\pm i\gamma B t/2}\ket{z\pm}
\]

Looking at the state $\ket{\psi(0)}$, we can write it as a superposition of the eigenstates of $H$
\[
	\ket{\psi(0)}=\cos\theta_0/2\ket{z+}+e^{i\phi_0}\sin\theta_0/2\ket{z-}
\]
\[
	\ket{\psi(t)}=U(t)\ket{\psi(0)}
\]
\[
	\ket{\psi(t)}=(e^{i\gamma B t/2})\cos\theta_0/2\ket{z+}+(e^{-i\gamma B t/2})e^{i\phi_0}\sin\theta_0/2\ket{z-}
\]
Global phase should be separated out
\[
	\ket{\psi(t)}=e^{i\gamma B t/2}\left[\cos\theta_0/2\ket{z+}+(e^{i(-\gamma B t+\phi_0)})\sin\theta_0/2\ket{z-}\right]
\]
Hence
\[
	\langle\psi(t)|\hat{S}_z|\psi(t)\rangle = \frac{\hbar}{2}\begin{pmatrix}\psi_+^* & \psi_-^*\end{pmatrix}\begin{pmatrix}1 & 0 \\ 0 & -1\end{pmatrix}\begin{pmatrix}\psi_+ \\ \psi_-\end{pmatrix}
\]

\[
	= \frac{\hbar}{2}\begin{pmatrix}\psi_+^* & \psi_-^*\end{pmatrix}\begin{pmatrix}\psi_+ \\ -\psi_-\end{pmatrix}
\]

\[
	= \frac{\hbar}{2}\left(\psi_+^*\psi_+ - \psi_-^*\psi_-\right) = \frac{\hbar}{2}\left(|\psi_+|^2 - |\psi_-|^2\right)
\]

\[
	= \frac{\hbar}{2}\left(\cos^2\frac{\theta_0}{2} - \sin^2\frac{\theta_0}{2}\right)
\]

\[
	= \frac{\hbar}{2}\cos\theta_0 \quad \text{(Note: unchanging in time)}
\]

\[
	\langle\psi(t)|\hat{S}_x|\psi(t)\rangle = \frac{\hbar}{2}\begin{pmatrix}\psi_+^* & \psi_-^*\end{pmatrix}\begin{pmatrix}0 & 1 \\ 1 & 0\end{pmatrix}\begin{pmatrix}\psi_+ \\ \psi_-\end{pmatrix}
\]

\[
	= \frac{\hbar}{2}\begin{pmatrix}\psi_+^* & \psi_-^*\end{pmatrix}\begin{pmatrix}\psi_- \\ \psi_+\end{pmatrix}
\]

\[
	= \frac{\hbar}{2}\left(\psi_+^*\psi_- + \psi_-^*\psi_+\right)
\]

Plug in \(\psi_+ = e^{+i\gamma B t/2}\cos(\theta_0/2)\) and \(\psi_- = e^{-i\gamma B t/2}e^{i\phi_0}\sin(\theta_0/2)\):

\[
	\psi_+^*\psi_- = e^{-i\gamma B t/2}\cos\frac{\theta_0}{2} \cdot e^{-i\gamma B t/2}e^{i\phi_0}\sin\frac{\theta_0}{2} = e^{i(\phi_0 - \gamma B t)}\cos\frac{\theta_0}{2}\sin\frac{\theta_0}{2}
\]

\[
	\psi_-^*\psi_+ = e^{-i(\phi_0 - \gamma B t)}\cos\frac{\theta_0}{2}\sin\frac{\theta_0}{2}
\]

\[
	\psi_+^*\psi_- + \psi_-^*\psi_+ = 2\cos\frac{\theta_0}{2}\sin\frac{\theta_0}{2}\cos(\phi_0 - \gamma B t) = \sin\theta_0 \cos(\phi_0 - \gamma B t)
\]

\[
	\Rarr \langle\psi(t)|\hat{S}_x|\psi(t)\rangle = \frac{\hbar}{2}\sin\theta_0 \cos(\phi_0 - \gamma B t)
\]

\[
	\langle\psi(t)|\hat{S}_y|\psi(t)\rangle = \frac{\hbar}{2}\begin{pmatrix}\psi_+^* & \psi_-^*\end{pmatrix}\begin{pmatrix}0 & -i \\ i & 0\end{pmatrix}\begin{pmatrix}\psi_+ \\ \psi_-\end{pmatrix}
\]

\[
	= \frac{\hbar}{2}\begin{pmatrix}\psi_+^* & \psi_-^*\end{pmatrix}\begin{pmatrix}-i\psi_- \\ i\psi_+\end{pmatrix}
\]

\[
	= \frac{\hbar}{2}\left(-i\psi_+^*\psi_- + i\psi_-^*\psi_+\right) = \frac{i\hbar}{2}\left(\psi_-^*\psi_+ - \psi_+^*\psi_-\right)
\]

Using the same products as above, with \(e^{ix} - e^{-ix} = 2i\sin x\):

\[
	\psi_-^*\psi_+ - \psi_+^*\psi_- = -2i\cos\frac{\theta_0}{2}\sin\frac{\theta_0}{2}\sin(\phi_0 - \gamma B t) = -i\sin\theta_0\sin(\phi_0 - \gamma B t)
\]

\[
	\Rarr \langle\psi(t)|\hat{S}_y|\psi(t)\rangle = \frac{i\hbar}{2}\cdot(-i)\sin\theta_0\sin(\phi_0 - \gamma B t) = \frac{\hbar}{2}\sin\theta_0\sin(\phi_0 - \gamma B t)
\]

Another way to do it is to use the \hyperref[heisenberg-picture]{Heisenberg picture}

\[
	\frac{d}{dt}\hat{A}(t) = \frac{i}{\hbar}[\hat{H}, \hat{A}]
\]

Recalling the \hyperref[pauli-matrices]{Pauli matrices} and their commutators, we can derive it for \hyperref[quantum-system]{spin operators} $S_z,S_x,S_y$.

\[
	[\hat{S}_x, \hat{S}_y] = i\hbar\hat{S}_z, \quad [\hat{S}_y, \hat{S}_z] = i\hbar\hat{S}_x, \quad [\hat{S}_z, \hat{S}_x] = i\hbar\hat{S}_y
\]

For our Hamiltonian \(\hat{H} = -\gamma B \hat{S}_z\):

\[
	\frac{d}{dt}\hat{S}_z(t) = \frac{i}{\hbar}[-\gamma B \hat{S}_z(t), \hat{S}_z(t)] = -\frac{i\gamma B}{\hbar}[\hat{S}_z(t), \hat{S}_z(t)] = 0
\]

\[
	\frac{d}{dt}\hat{S}_x(t) = \frac{i}{\hbar}[-\gamma B \hat{S}_z(t), \hat{S}_x(t)]
\]

\[
	= -\frac{i\gamma B}{\hbar}[\hat{S}_z(t), \hat{S}_x(t)]
\]

\[
	= -\frac{i\gamma B}{\hbar}\cdot i\hbar \hat{S}_y(t) = \gamma B \hat{S}_y(t)
\]

\[
	\frac{d}{dt}\hat{S}_y(t) = \frac{i}{\hbar}[-\gamma B \hat{S}_z(t), \hat{S}_y(t)]
\]

\[
	= -\frac{i\gamma B}{\hbar}[\hat{S}_z(t), \hat{S}_y(t)]
\]

\[
	= -\frac{i\gamma B}{\hbar}\cdot(-i\hbar \hat{S}_x(t)) = -\gamma B \hat{S}_x(t)
\]

Collecting all three:

\[
	\frac{d}{dt}(\hat{S}_x, \hat{S}_y, \hat{S}_z) = \gamma B(\hat{S}_y, -\hat{S}_x, 0)
\]

Using \hyperref[cross-product]{cross product} formula:

\[
	= -\gamma(0, 0, B)\times(\hat{S}_x, \hat{S}_y, \hat{S}_z)
\]

\[
	\boxed{\frac{d\vec{S}}{dt} = -\gamma \vec{B}\times\vec{S}} \quad \text{(Larmor precession)}
\]

\section{General Qubit Hamiltonian and Dynamics}\label{general-qubit-hamiltonian-and-dynamics}
Relies on \hyperref[quantum-dynamics]{quantum dynamics}.

This means that any 2x2 Hermitian Hamiltonian can be decomposed into an identity plus a rotation generator part.

\[ H = \begin{pmatrix} g_0 + g_3 & g_1 - ig_2 \\ g_1 + ig_2 & g_0 - g_3 \end{pmatrix} \quad g_i \in \mathbb{R} \]

\[ = g_0 I + g_1 X + g_2 Y + g_3 Z \]

\[ = g_0 I + \vec{g}\cdot\vec{\sigma} \qquad \vec{g} = (g_1, g_2, g_3) = g\hat{n} \]

\[ = g_0 I + g\,\hat{n}\cdot\vec{\sigma} \qquad \vec{\sigma} = (X, Y, Z) = (\sigma_x, \sigma_y, \sigma_z) \]

\[ = g_0 I + \frac{2g}{\hbar}\,\hat{n}\cdot\vec{S} \]


Remarks
\begin{enumerate}
	\item If and only $B$ is Hermitian then $U=e^{iB}$ is unitary.
	\item Spin/angular momentum generates rotations
	\item Unitary $V$ is a dynamical symmetry of system if it commutes with time evolution
	      \[
		      [V,U(t)]=0 \Rarr [V,H(t)]=0
	      \]
	\item Hermitian $A$ is a conserved quantity if $[A,H(t)]=0$. This also means $V(\theta)=e^{i\theta A}$ is a dynamical symmetry of the system. A conserved quantity generates dynamical symmetries.

\end{enumerate}
\section{Time Energy Uncertainty}\label{time-energy-uncertainty}
Relies on \hyperref[quantum-dynamics]{quantum dynamics}.

Time is not an observable.

We could count $N$ full waves of frequency $\omega$ over a time window $T$. You can only count to within $\pm 1$ wave, so $\Delta N\geq 1$.

The uncertainty in the wavecount is $\Delta N$

We know that

\[
	N=\frac{\omega T}{2\pi} \quad\Rarr \quad \Delta N = \frac{\Delta \omega T}{2\pi} \geq 1 \quad \Rarr \quad \Delta \omega T \geq 2\pi
\]

From \hyperref[planck-einstein-relation](Planck-Einstein relation),
\[
	\Delta E \cdot T\geq h = \frac{\hbar}{2}
\]
\[
	\begin{pmatrix}
		\alpha   & \beta    \\
		-\beta^* & \alpha^*
	\end{pmatrix}
\]
\end{document}

